--- 
title_meta  : Chapter 4
title       : R Functionals
description : See details in R function settings


--- type:NormalExercise lang:r xp:100 skills:1 key:36ba783482
## Lazy Evaluation of Function Arguments

R functions delay evaluation of their arguments until they’re needed by their R code, This is called lazy evaluation, If the function body doesn’t evaluate an argument, then the function won’t produce an error, even if the argument is missing.


*** =instructions
- Input varied arguments to R functions

*** =hint
Follow the instructions and introductions.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# define function lazy_func
lazy_func <- function(arg1, arg2) {
  2*arg1
}

# bind arguments by position


# define function lazy_func
lazy_func <- function(arg1, arg2) {
  cat(arg1, '\n')
  cat(arg2)
}

# bind arguments by position


# first argument written to output, without second argument

```

*** =solution
```{r}
# define function lazy_func
lazy_func <- function(arg1, arg2) {
  2*arg1
}

# bind arguments by position
lazy_func(3, 2)

# define function lazy_func
lazy_func <- function(arg1, arg2) {
  cat(arg1, '\n')
  cat(arg2)
}

# bind arguments by position
lazy_func(3, 2)

# first argument written to output, without the second argument
lazy_func(3)
```

*** =sct
```{r}
success_msg("Great!");
```


--- type:NormalExercise lang:r xp:100 skills:1 key:56d9d5409e
## The dots "..." Function Argument

The dots `...` function argument is a formal argument without a name, as opposed to the other formal arguments which all have names, The dots `...` bind with any number of additional arguments, that aren’t already bound by name or position to the named arguments, The dots "..." are used when the number of arguments isn’t known in advance, and allows functions to accept an indeﬁnite number of arguments, The dots `...` are sometimes placed after the named arguments, to allow passing of additional parameters into a function, Functionals often place the dots `...` argument after the named arguments, to allow passing the dots `...` to the function being called by the functional.

*** =instructions
- Learn to use `...` for flexible length of argument

*** =hint
Follow the instruction and introduction.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# dots for additional plot parameters


# change the code to create a function that accepts undefined length of arguments
bind_dots <- function(in_put, ...) {
  paste0("in_put=", in_put,
 ", dots=", paste(..., sep=", "))
}

# "in_put" bound by position


# "in_put" bound by name


# named argument bound to dots


# define a new function with at least two argument for input
bind_dots <- function(arg1, arg2, ...) {
  arg1 + 2*arg2 + sum(...)
}

# bind arguments by position
bind_dots(3, 2)

# extra arguments bound to dots


# dots before other arguments


# dots bind before other arguments


# change the default argument


# define a new function with default argument at the end
bind_dots <- function(..., in_put=10) {
  paste0("in_put=", in_put,
 ", dots=", paste(..., sep=", "))
}

# arguments after dots must be bound by full name


# dots bound


# "in_put" not bound, but has default

```

*** =solution
```{r}
# dots for additional plot parameters
str(plot)

# change the code to create a function that accepts undefined length of arguments
bind_dots <- function(in_put, ...) {
  paste0("in_put=", in_put,
 ", dots=", paste(..., sep=", "))
}

# "in_put" bound by position
bind_dots(1, 2, 3)

# "in_put" bound by name
bind_dots(2, in_put=1, 3)

# named argument bound to dots
bind_dots(1, 2, 3, foo=10)

# define a new function with at least two argument for input
bind_dots <- function(arg1, arg2, ...) {
  arg1 + 2*arg2 + sum(...)
}

# bind arguments by position
bind_dots(3, 2)

# extra arguments bound to dots
bind_dots(3, 2, 5, 8)

# dots before other arguments
str(sum)

# dots bind before other arguments
sum(1, 2, 3)

# change the default argument
sum(1, 2, NA, 3, na.rm=TRUE)

# define a new function with default argument at the end
bind_dots <- function(..., in_put=10) {
  paste0("in_put=", in_put,
 ", dots=", paste(..., sep=", "))
}

# arguments after dots must be bound by full name
bind_dots(1, 2, 3, in_put=10)

# dots bound
bind_dots(1, 2, 3, in_put=10, foo=4)

# "in_put" not bound, but has default
bind_dots(1, 2, 3)
```

*** =sct
```{r}
success_msg("Wonderful! Proceed to the next exercise.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:e09996c475
## Functions that takes functions and returns functions

Functions in R are ﬁrst class objects, which means they can be treated like any other R object: Functions can be passed as arguments to other functions, Functions can be nested (deﬁned inside other functions), Functions can return functions as their return value, Higher-order functions are R functions that either accept a function as their argument (input) or return a function as their value (output).

R functions can also return a function as their value, Functions returned by a function are called closures, Functions that return closures can be used as function factories.

*** =instructions
- Use function name as input, and get returns as some functions.

*** =hint
Follow the instructions and introductions.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# create functional that accepts a function as input argument
func_tional <- function(func_name) {

# set seed to random numbers
  set.seed(1)

# change the code to apply the function name
  func(runif(1e4))
}

# get mean
func_tional(mean)

# get standard deviation
func_tional(sd)

# define a power function factory
make_func <- function(arg_param) {
  function(in_put) {
    in_put^arg_param
  }
}

# define square function
square_func <- make_func(2)

# input 4
square_func(4)

# define cube function
cube_func <- make_func(3)

# apply 2
cube_func(2)

# define cube root function


# apply 8

```

*** =solution
```{r}
# create functional that accepts a function as input argument
func_tional <- function(func_name) {

# set seed to random numbers
  set.seed(1)

# change the code to apply the function name
  func(runif(1e4))
}

# get mean


# get standard deviation


# define a power function factory
make_func <- function(arg_param) {
  function(in_put) {
    in_put^arg_param
  }
}

# define square function


# input 4


# define cube function


# apply 2


# define cube root function


# apply 8

```

*** =sct
```{r}
success_msg("Awesome! You correctly figured them out.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:f6eb0a5c5f
## Functionals

Functionals are functions that accept a function or a function name as one of their input arguments, Functionals are able to execute function calls using the function names, The function match.fun() returns a function name that is speciﬁed by a string, match.fun() produces an error condition if it fails to ﬁnd a function with the speciﬁed name.

The dots "..." argument in functionals can be used to pass additional arguments to the function being called by the functional, If named values are passed to the dots "..." argument, then the functional can bind them to the correct formal arguments of the function being called by the functional.

*** =instructions
- Use function name and dots for function inputs. 

*** =hint
Follow the instruction and introductions.
Use `match.fun()` to convert function name to function.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# func_tional accepts function name and additional argument
func_tional <- function(func_name, in_put) {

# change the code to produce function name from argument
  func_name <- func_name

# change the code to execute function call
  func_name()
}

# get the square root of 4


# sum() accepts multiple arguments
# func_tional can't accept indefinite number of arguments


# func_tional accepts function name and dots '...' argument
func_tional <- function(func_name, ...) {
  func_name <- match.fun(func_name)
  func_name(...)
}

# get the sum of 1, 2, 3


# get the sum containing NAs


# get the sum removing NAs


# function with three arguments and dots '...' arguments
my_func <- function(in_put, param1, param2, ...) {
  c(input=in_put, param1=param1, param2=param2,
dots=c(...))
}

# get a named vector

```

*** =solution
```{r}
# func_tional accepts function name and additional argument
func_tional <- function(func_name, in_put) {

# change the code to produce function name from argument
  func_name <- match.fun(func_name)

# change the code to execute function call
  func_name(in_put)
}

# get the square root of 4
func_tional(sqrt, 4)

# sum() accepts multiple arguments
# func_tional can't accept indefinite number of arguments
str(sum)

# func_tional accepts function name and dots '...' argument
func_tional <- function(func_name, ...) {
  func_name <- match.fun(func_name)
  func_name(...)
}

# get the sum of 1, 2, 3
func_tional(sum, 1, 2, 3)

# get the sum containing NAs
func_tional(sum, 1, 2, NA, 4, 5)

# get the sum removing NAs
func_tional(sum, 1, 2, NA, 4, 5, na.rm=TRUE)

# function with three arguments and dots '...' arguments
my_func <- function(in_put, param1, param2, ...) {
  c(input=in_put, param1=param1, param2=param2,
dots=c(...))
}

# get a named vector
my_func(1, 2, 3, param2=4, param1=5)
```

*** =sct
```{r}
success_msg("Phew, it seems that R is flexible enough ;-).")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:d65e71390a
## Anonymous Functions and Functionals

R allows deﬁning functions without assigning a name to them, Anonymous functions are functions that are not assigned to a name, Anonymous functions can be passed as arguments to functionals.

Anonymous functions can be passed as arguments to functionals, Anonymous functions can also be used as default values for function arguments.

*** =instructions
- Produce compact functionality with the anonymous functions and apply them to functionals.

*** =hint
Remember to write anonymous functions with complete structure.

*** =pre_exercise_code
```{r}
# no pec
func_tional <- function(func_name, ...) {
  func_name <- match.fun(func_name)
  func_name(...)
}
```

*** =sample_code
```{r}
# simple anonymous function
(function(x) (x + 3)) (10)

# anonymous function passed to func_tional
func_tional(func_name=(function(x) (x + 3)), 5)

# change the code to make anonymous function is default value
func_tional <-
  function(..., func_name=function(x, y, z) {x+y+z}) {
    func_name <- func_name
    func_name()
}

# use default func_name
# more arguments result an error
func_tional(2, 3, 4)

# func_name bound by name
func_tional(func_name=sum, 2, 3, 4, 5)

# change the code to pass anonymous function to func_name
func_tional(func_name=function(x, y, z),
    2, 3, 4)
```

*** =solution
```{r}
# simple anonymous function
(function(x) (x + 3)) (10)

# anonymous function passed to func_tional
func_tional(func_name=(function(x) (x + 3)), 5)

# change the code to make anonymous function is default value
func_tional <-
  function(..., func_name=function(x, y, z) {x+y+z}) {
    func_name <- match.fun(func_name)
    func_name(...)
}

# use default func_name
# more arguments result an error
func_tional(2, 3, 4)

# func_name bound by name
func_tional(func_name=sum, 2, 3, 4, 5)

# pass anonymous function to func_name
func_tional(func_name=function(x, y, z) {x*y*z},
    2, 3, 4)
```

*** =sct
```{r}
success_msg("A job well done! Continue to the next exercise.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:e23011c42b
## Executing Function Calls Using the do.call() Functional

The functional do.call() executes a function call using a function name and a list of arguments, do.call() allows calling a function on arguments that are elements of a list, do.call() passes the list elements individually, instead of passing the whole list as one argument: do.call(fun, list)= fun(list[[1]], list[[2]], ...)

do.call() can be called inside other functionals to allow them to execute function calls, The function str() displays the structure of an R object, for example a function name and its formal arguments.

*** =instructions
Use `do.call()` to execute function.

*** =hint
Always call function name as the first argument and inputs as following arguments.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# "what" argument is a function


# do.call passes list elements into "sum" individually


# passes list element containing NA


# filter NAs in list elements


# func_tional() accepts list with function name and arguments
func_tional <- function(list_arg) {

# change the code to produce function name from argument
  func_name <- match.fun(list_arg)
  
# change the code to execute function call uing do.call()
  do.call(func_name, list_arg)
}

# construction an arg_list of function name and function input


# pass the argument to functional

```

*** =solution
```{r}
# "what" argument is a function
str(do.call)

# do.call passes list elements into "sum" individually
do.call(sum, list(1, 2, 3))

# passes list element containing NA
do.call(sum, list(1, 2, NA, 3))

# filter NAs in list elements
do.call(sum, list(1, 2, NA, 3, na.rm=TRUE))

# func_tional() accepts list with function name and arguments
func_tional <- function(list_arg) {

# produce function name from argument
  func_name <- match.fun(list_arg[[1]])
  
# execute function call uing do.call()
  do.call(func_name, list_arg[-1])
}

# construction an argument list of function name and function input
arg_list <- list("sum", 1, 2, 3)

# pass the argument to functional
func_tional(arg_list)
```

*** =sct
```{r}
success_msg("Bellissimo!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:59a81e117a
## Performing Loops Using the apply() Functionals

An important example of functionals are the apply() functions, The function apply() returns the result of applying a function to the rows or columns of an array or matrix, If MARGIN=1 then the function will be applied over the matrix rows, If MARGIN=2 then the function will be applied over the matrix columns, apply() performs a loop over the list of objects, and can replace "for" loops in R.


*** =instructions
Use `apply()` to execute function in loops.

*** =hint
Always call data object as the first argument of apply, second arguemnt is the dimension, and function name as following arguments.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# get list of arguments


# create a matrix
mat_rix <- matrix(6:1, nrow=2, ncol=3)

# print the matrix


# sum the rows


# sum the columns


# change the code define a matrix of sums and original matrix
mat_rix <- c(sum(row_sums), col_sums)
dimnames(mat_rix) <- list(c("col_sums", "row1", "row2"),
                 c("row_sums", "col1", "col2", "col3"))

# print the matrix

```

*** =solution
```{r}
# get list of arguments
str(apply)

# create a matrix
mat_rix <- matrix(6:1, nrow=2, ncol=3)

# print the matrix
mat_rix

# sum the rows
row_sums <- apply(mat_rix, 1, sum)

# sum the columns
col_sums <- apply(mat_rix, 2, sum)

# change the code define a matrix of sums and original matrix
mat_rix <- cbind(c(sum(row_sums), row_sums),
          rbind(col_sums, mat_rix))
dimnames(mat_rix) <- list(c("col_sums", "row1", "row2"),
                 c("row_sums", "col1", "col2", "col3"))

# print the matrix
mat_rix
```

*** =sct
```{r}
success_msg("Groß!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:8debcc371a
## The apply() Functional with dots "..." Argument

The dots "..." argument in apply() is designed to pass additional arguments to the function being called by apply(), The additional arguments to apply() must be bound by their full (complete) names.


*** =instructions
Use `apply()` to execute function in loops with additional arguments.

*** =hint
Always call data object as the first argument of apply, second arguemnt is the dimension, and function name as following arguments. More arguments can be passed as dot arguments.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# get list of arguments


# create a matrix
mat_rix <- matrix(sample(12), nrow=3, ncol=4)

# print the code


# sort matrix columns


# sort decreasing order


# introduce NA value


# print the matrix


# calculate median of columns


# calculate median of columns with na.rm=TRUE

```

*** =solution
```{r}
# get list of arguments
str(apply)

# create a matrix
mat_rix <- matrix(sample(12), nrow=3, ncol=4)

# print the code
mat_rix

# sort matrix columns
apply(mat_rix, 2, sort)

# sort decreasing order
apply(mat_rix, 2, sort, decreasing=TRUE)

# introduce NA value
mat_rix[2, 2] <- NA

# print the matrix
mat_rix

# calculate median of columns
apply(mat_rix, 2, median)

# calculate median of columns with na.rm=TRUE
apply(mat_rix, 2, median, na.rm=TRUE)
```

*** =sct
```{r}
success_msg("Doskonały!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:2ce2587280
## The apply() Functional with Anonymous Functions

The apply() functional combined with anonymous functions can be used to loop over function parameters, The dots "..." argument in apply() is designed to pass additional arguments to the function being called by apply(), The additional arguments to apply() must be bound by their full (complete) names.


*** =instructions
Use `apply()` to execute your customized function in loops.

*** =hint
Always call data object as the first argument of apply, second arguemnt is the dimension, and function name as following arguments. Your customized function could be defined in the third argument.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# DAX percent returns
dax_rets

# load package moments


# get list of arguments of moment


# apply moment function


# 4x1 matrix of moment orders
moment_orders <- as.matrix(1:4)

# change the code so that anonymous function allows looping over function parameters
apply(X=moment_orders, MARGIN=1,
      FUN=function(moment_order) {

}
      )  # end apply

# another way of passing parameters into moment() function
apply(X=moment_orders, MARGIN=1, FUN=moment)
```

*** =solution
```{r}
# DAX percent returns
dax_rets <- 100*diff(log(EuStockMarkets[, 1]))

# load package moments
library(moments)

# get list of arguments of moment
str(moment)

# apply moment function
moment(x=dax_rets, order=3)

# 4x1 matrix of moment orders
moment_orders <- as.matrix(1:4)

# change the code so that anonymous function allows looping over function parameters
apply(X=moment_orders, MARGIN=1,
      FUN=function(moment_order) {
  moment(x=dax_rets, order=moment_order)
}
      )  # end apply

# another way of passing parameters into moment() function
apply(X=moment_orders, MARGIN=1, FUN=moment,
      x=dax_rets)
```

*** =sct
```{r}
success_msg("Strålende!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:ea0beb2602
## apply() Calling Functions with Multiple Arguments

When apply() calls a function with multiple arguments, then care must be taken for proper argument binding, The dots "..." argument in apply() allows passing additional arguments to the function being called by apply(), The additional arguments to apply() must be bound by their full (complete) names, The values of the "X" argument in apply() are bound by position to the ﬁrst unused argument in the function being called by apply().

*** =instructions
Use `apply()` to execute function in loops.

*** =hint
Always call data object as the first argument of apply, second arguemnt is the dimension, and function name as following arguments. 

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# function with three arguments
my_func <- function(arg1, arg2, arg3) {
  c(arg1=arg1, arg2=arg2, arg3=arg3)
}

# input 1,2,3 to function


# define a matrix


# pass da_ta to arg1


# pass da_ta to arg2


# pass da_ta to arg3

```

*** =solution
```{r}
# function with three arguments
my_func <- function(arg1, arg2, arg3) {
  c(arg1=arg1, arg2=arg2, arg3=arg3)
}

# input 1,2,3 to function
my_func(1, 2, 3)

# define a matrix
da_ta <- as.matrix(1:4)

# pass da_ta to arg1
apply(X=da_ta, MAR=1, FUN=my_func, arg2=2, arg3=3)

# pass da_ta to arg2
apply(X=da_ta, MAR=1, FUN=my_func, arg1=1, arg3=3)

# pass da_ta to arg3
apply(X=da_ta, MAR=1, FUN=my_func, arg1=1, arg2=2)
```

*** =sct
```{r}
success_msg("Apply loop over a n*1 matrix and returns a full matrix!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:ce02dc9d63
## The lapply() and sapply() Functional

When apply() calls a function with multiple arguments, then care must be taken for proper argument binding, The dots "..." argument in apply() allows passing additional arguments to the function being called by apply(), The additional arguments to apply() must be bound by their full (complete) names, The values of the "X" argument in apply() are bound by position to the ﬁrst unused argument in the function being called by apply().

*** =instructions
Use `lapply()` and `sapply()` to execute function in loops.

*** =hint
Lapply accepts a list and returns a list, while sapply() tries to return a matrix.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# vector of means of numeric columns


# list of means of numeric columns


# change the code to fix bugs and let lapply use anonymous function
unlist(lapply(iris,
      function(col_umn) {
        mean(col_umn)
      }
      )
       )

# change the code to get flattened results
sapply(iris, function(col_umn) {
  mean(col_umn)})

# sapply on vector


# sapply on list


# calculate means of iris data frame columns


# create a matrix


# calculate column means using apply

```

*** =solution
```{r}
# vector of means of numeric columns
sapply(iris[, -5], mean)

# list of means of numeric columns
lapply(iris[, -5], mean)

# change the code to fix bugs and let lapply use anonymous function
unlist(lapply(iris,
      function(col_umn) {
        if (is.numeric(col_umn)) mean(col_umn)
      }
      )
       )

# flatten the list or matrix
unlist(sapply(iris, function(col_umn) {
  if (is.numeric(col_umn)) mean(col_umn)}))

# sapply on vector
sapply(6:10, sqrt)

# sapply on list
sapply(list(6, 7, 8, 9, 10), sqrt)

# calculate means of iris data frame columns
sapply(iris, mean)

# create a matrix
mat_rix <- matrix(sample(100), ncol=4)

# calculate column means using apply
apply(mat_rix, 2, mean)
```

*** =sct
```{r}
success_msg("You've just learned to use two powerful weapons in the world of R")
```