--- 
title_meta  : Chapter 5
title       : More on matrix
description : Matrix, and matrix-like structures are essential to R data manipulation and operation.

--- type:NormalExercise lang:r xp:100 skills:1 key:b14f9472dd
## Logical Operators Applied to Vectors and Matrices

When logical operators are applied to vectors and matrices, they are applied element-wise, producing boolean vectors and matrices.

*** =instructions
- Apply logical operators to matrix and vectors.

*** =hint
Follow the instruction and introduction.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# create a vector
vec_tor <- sample(1:9)

# print the vector


# element-wise comparison if it's more than 5


# element-wise comparison if it's equal to 5


# create a matrix
mat_rix <- matrix(vec_tor, ncol=3)

# print the matrix


# element-wise comparison if it's less than 5


# element-wise comparison if it's equal to 5

```

*** =solution
```{r}
# create a vector
vec_tor <- sample(1:9)

# print the vector
vec_tor

# element-wise comparison if it's more than 5
vec_tor < 5

# element-wise comparison if it's equal to 5
vec_tor == 5

# create a matrix
mat_rix <- matrix(vec_tor, ncol=3)

# print the matrix
mat_rix

# element-wise comparison if it's less than 5
mat_rix < 5

# element-wise comparison if it's equal to 5
mat_rix == 5
```

*** =sct
```{r}
test_error()
success_msg("Wonderful! Your skillset is growing at a staggering pace! Head over to the next exercise.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:849f04f218
## Coercing Vectors Into Matrices

Vectors can be coerced into matrices by adding a dimension attribute, The dimnames attribute can be assigned a named list to convert it into a named matrix, The function structure() adds attributes (speciﬁed as symbol=value pairs) to an object, and returns it.


*** =instructions
- Learn to coerce vectors and create matrix.

*** =hint
Follow the instruction and introduction.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# create a vector
mat_rix <- 1:6

# get its class


# is it vector or matrix?


# matrix object


# adding dimension attribute coerces into matrix


# get its class


# is it vector or matrix?


# assign dimnames attribute

                  
# print the matrix

```

*** =solution
```{r}
# create a vector
mat_rix <- 1:6

# get its class
class(mat_rix)

# is it vector or matrix?
c(is.vector(mat_rix), is.matrix(mat_rix))

# matrix object
structure(mat_rix, dim=c(2, 3))

# adding dimension attribute coerces into matrix
dim(mat_rix) <- c(2, 3)

# get its class
class(mat_rix)

# is it vector or matrix?
c(is.vector(mat_rix), is.matrix(mat_rix))

# assign dimnames attribute
dimnames(mat_rix) <- list(rows=c("row1", "row2"),
                  columns=c("col1", "col2", "col3"))
                  
# print the matrix
mat_rix
```

*** =sct
```{r}
test_error()
success_msg("Nice one! Can you see from the displayed structure of matrix. Next!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:f37117a766
## Coercing Matrices Into Other Types

Matrices can be explicitly coerced using the `as.*` coercion functions, But coercion functions strip the attributes from an object.


*** =instructions
- Use `as.numeric`, `as.character` and other similar functions to coerce matrix to other types.

*** =hint
Follow the instruction and introduction.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# create matrix


# print the matrix


# as.numeric strips dim attribute from matrix


# explicitly coerce to "character"


# check the type, mode and class of new variable


# coercion converted matrix to vector

```

*** =solution
```{r}
# create matrix
mat_rix <- matrix(1:10, 2, 5)

# print the matrix
mat_rix

# as.numeric strips dim attribute from matrix
as.numeric(mat_rix)

# explicitly coerce to "character"
mat_rix <- as.character(mat_rix)

# check the type, mode and class of new variable
c(typeof(mat_rix), mode(mat_rix), class(mat_rix))

# coercion converted matrix to vector
c(is.matrix(mat_rix), is.vector(mat_rix))
```

*** =sct
```{r}
test_error()
success_msg("Great!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:dc5f4f6f30
## Binding Vectors and Matrices Together

Vectors can be bound into matrices using the functions cbind() and rbind(), The recycling rule allows operations on vectors of diﬀerent lengths:

1 Vectors are scanned from left to right, 

2 Shorter vectors are extended in length by recycling their values until they match the length of longer vectors.


*** =instructions
- Create matrices by binding vectors, even for different lengths!

*** =hint
Follow the instruction and introduction.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# define vector
vec_tor1 <- 1:3

# define vector
vec_tor2 <- 6:4

# bind into columns


# bind into rows


# extend to four elements


# recycling rule applied


# another example of recycling rule

```

*** =solution
```{r}
# define vector
vec_tor1 <- 1:3

# define vector
vec_tor2 <- 6:4

# bind into columns
cbind(vec_tor1, vec_tor2)

# bind into rows
rbind(vec_tor1, vec_tor2)

# extend to four elements
vec_tor2 <- c(vec_tor2, 7)

# recycling rule applied
cbind(vec_tor1, vec_tor2)

# another example of recycling rule
1:6 + c(10, 20)
```

*** =sct
```{r}
test_error()
success_msg("Perfect!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:1e3b5f4b0a
## Replicating Objects Using rep()

The function rep() replicates vectors and lists a given number of times, rep() accepts a vector or list "x", and an integer specifying the type and number of replications.

Argument "times" replicates the whole vector a given number of times.

Argument "each" replicates each vector element a given number of times.

Argument "length.out" replicates the whole vector a certain number of times, so that the output vector length is equal to "length.out".


*** =instructions
- Change the arguments in the introduction to manipulate output of `rep()`

*** =hint
Follow the instruction and introduction.

*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
# replicate a single element


# replicate the whole vector several times


# explicitly set the times


# replicate the first element, then the second, etc.


# replicate to specified length

```

*** =solution
```{r}
# replicate a single element
rep("a", 5)

# replicate the whole vector several times
rep(c("a", "b"), 5)

# explicitly set the times
rep(c("a", "b"), times=5)

# replicate the first element, then the second, etc.
rep(c("a", "b"), each=5)

# replicate to specified length
rep(c("a", "b"), length.out=5)
```

*** =sct
```{r}
test_error()
success_msg("Nice! That was still pretty easy, right?")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:cbda2853c3
## Multiplying Vectors and Matrices

The multiplication `*` operator performs element-wise (element-by-element) multiplication of vectors and matrices.

By default the matrix elements are multiplied column-wise by the vector elements: the ﬁrst matrix element in the ﬁrst column is multiplied by the ﬁrst vector element, then the second matrix column is multiplied by the remaining vector elements, etc. 

The recycling rule is applied to the vector elements as needed, The transpose function `t()` can be applied if we want to perform row-wise multiplication.


*** =instructions
- Use `*` for element-wise multiplication.

*** =hint
Follow the instruction and introduction.

*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
# define a vector
vec_tor1 <- c(2, 4, 3)

# define a matrix
mat_rix <- matrix(sample(1:12), ncol=3)

# multiply matrix by vector column-wise


# multiply matrix by vector row-wise

```

*** =solution
```{r}
# define a vector
vec_tor1 <- c(2, 4, 3)

# define a matrix
mat_rix <- matrix(sample(1:12), ncol=3)

# multiply matrix by vector column-wise
vec_tor1 * mat_rix

# multiply matrix by vector row-wise
t(vec_tor1 * t(mat_rix))
```

*** =sct
```{r}
test_error()
success_msg("Great! Multiply by elements is rather easy isn't it? Continue to the next exercise.")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:35ba0d6f0d
## Matrix Inner Multiplication

The `%*%` operator performs inner (scalar) multiplication of vectors and matrices, Inner multiplication consists of multiplying the rows of one matrix with the columns of another matrix, so that each pair produces a single number.

Inner multiplication produces a vector or matrix with a reduced dimension, Inner multiplication requires the dimensions of the matrices to be conformable (number of columns in the ﬁrst matrix must be equal to the number of rows in the second), otherwise an error is produced, The function `drop()` removes any dimensions of length one.

*** =instructions
- Use `%*%` for mathematical matrix multiplication.

*** =hint
Follow the instruction and introduction. Distinguish between `*` and `%*%``

*** =pre_exercise_code
```{r}
```

*** =sample_code

```{r}
# create matrix
mat_rix <- matrix(1:6, ncol=3)

# define the vector 1
vec_tor1 <- c(2, 4, 3)

# define vector
vec_tor2 <- 6:4

# multiply two vectors element-by-element


# calculate inner product


# calculate inner product and drop dimensions

```

*** =solution
```{r}
# create matrix
mat_rix <- matrix(1:6, ncol=3)

# define the vector 1
vec_tor1 <- c(2, 4, 3)

# define vector
vec_tor2 <- 6:4

# multiply two vectors element-by-element
vec_tor1 * vec_tor2

# calculate inner product
vec_tor1 %*% vec_tor2

# calculate inner product and drop dimensions
drop(vec_tor1 %*% vec_tor2)
```

*** =sct
```{r}
test_error()
success_msg("Congrats!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:bf49ce3efc
## Matrix Transpose

The function t() returns the transpose of a matrix, The function crossprod() also performs inner (scalar) multiplication, exactly the same as the %*% operator, but is slightly faster.

*** =instructions
- Use `t()` to transpose matrices.

*** =hint
Follow the instruction and introduction. 

*** =pre_exercise_code
```{r}
library(microbenchmark)
# create matrix
mat_rix <- matrix(1:6, ncol=3)

# define the vector 1
vec_tor1 <- c(2, 4, 3)

vec_tor2 <- 6:4
```

*** =sample_code

```{r}
# multiply matrix by vector fails because dimensions aren't conformable


# works after transpose


# calculate inner product


# create matrix and vector
mat_rix <- matrix(1:3000, ncol=3)

# transpose matrix


# create new vector
vec_tor <- 1:3

# change the code to see crossprod is slightly faster than "%*%" operator
summary(microbenchmark(
  cross_prod=prod(tmat_rix, vec_tor),
  inner_prod=mat_rix * vec_tor,
  times=10))[, c(1, 4, 5)]
```

*** =solution
```{r}
# multiply matrix by vector fails because dimensions aren't conformable
vec_tor1 %*% mat_rix

# works after transpose
drop(vec_tor1 %*% t(mat_rix))

# calculate inner product
crossprod(vec_tor1, vec_tor2)

# create matrix and vector
mat_rix <- matrix(1:3000, ncol=3)

# transpose matrix
tmat_rix <- t(mat_rix)

# create new vector
vec_tor <- 1:3

# change the code to see crossprod is slightly faster than "%*%" operator
summary(microbenchmark(
  cross_prod=crossprod(tmat_rix, vec_tor),
  inner_prod=mat_rix %*% vec_tor,
  times=10))[, c(1, 4, 5)]
```

*** =sct
```{r}
test_error()
success_msg("Takes a bit time, but learn from the result may save you a lot more time!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:b0b3b7fe27
## Matrix Outer Multiplication

An outer product consists of all possible products of pairs of elements of two objects: 
````
Ci,j = Ai ·Bj
````

An outer product of a function consists of applying it to all possible pairs of elements of two objects:
````
Ci,j = f (Ai,Bj)
````
Outer multiplication produces an object with dimension equal to the sum of the factors’ dimensions, and with the number of elements equal to the product of the factors’ elements, The function outer() calculates the outer product of two matrices, and by default multiplies the elements of its arguments, outer() can also calculate the values of a vectorized function of two variables passed to the "FUN" 

*** =instructions
- Use outer multiplicaiton to preserve and  extend dimension.

*** =hint
Follow the instruction and introduction. 

*** =pre_exercise_code
```{r}
```

*** =sample_code

```{r}
# define named vectors
vec_tor1 <- sample(1:4)

# assign names for rows


# print the vector


# create new vector
vec_tor2 <- sample(1:3)

# assign names for columns


# print the vector


# assign outer product of two vectors to mat_rix


# print the matrix

# change the code, calculate vectorized function spanned over two vectors
mat_rix <- outer(vec_tor1, vec_tor2,
           FUN=function() x2:sin(x1))

# print the matrix

```

*** =solution
```{r}
# define named vectors
vec_tor1 <- sample(1:4)

# assign names for rows
names(vec_tor1) <- paste0("row", 1:4, "=", vec_tor1)

# print the vector
vec_tor1

# create new vector
vec_tor2 <- sample(1:3)

# assign names for columns
names(vec_tor2) <- paste0("col", 1:3, "=", vec_tor2)

# print the vector
vec_tor2

# assign outer product of two vectors to mat_rix
mat_rix <- outer(vec_tor1, vec_tor2)

# print the matrix
mat_rix

# change the code, calculate vectorized function spanned over two vectors
mat_rix <- outer(vec_tor1, vec_tor2,
           FUN=function(x1, x2) x2*sin(x1))

# print the matrix
mat_rix
```

*** =sct
```{r}
test_error()
success_msg("That's potentially a new way to create matrix out of vectors")
```



--- type:NormalExercise lang:r xp:100 skills:1 key:eb77d79f8a
## Flattening a List of Vectors to a Matrix Using do.call()

A list of vectors can be ﬂattened into a matrix using the functions `do.call()` and either `rbind()` or `cbind()`, If the list contains vectors of diﬀerent lengths, then R applies the recycling rule, If the list contains a NULL element, that element is skipped,


*** =instructions
- Flatten list of vectors to more structured matrix.

*** =hint
Follow the instruction and introduction. 

*** =pre_exercise_code
```{r}
```

*** =sample_code

```{r}
# create list of vectors


# bind list elements into matrix - doesn't work


# bind list elements into matrix - tedious


# bind list elements into matrix - works!


# create numeric list
li_st <- list(1, 2, 3, 4)

# returns single column matrix


# returns single row matrix

```

*** =solution
```{r}
# create list of vectors
li_st <- lapply(1:3, function(x) sample(6))

# bind list elements into matrix - doesn't work
rbind(li_st)

# bind list elements into matrix - tedious
rbind(li_st[[1]], li_st[[2]], li_st[[3]])

# bind list elements into matrix - works!
do.call(rbind, li_st)

# create numeric list
li_st <- list(1, 2, 3, 4)

# returns single column matrix
do.call(rbind, li_st)

# returns single row matrix
do.call(cbind, li_st)
```

*** =sct
```{r}
test_error()
success_msg("Cool! Don't forget the do.call()!`")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:04f0d28b95
## Eﬃcient Binding of Lists Into Matrices

A list of vectors can be ﬂattened into a matrix using the functions `do.call()` and either `rbind()` or `cbind()`, If the list contains vectors of diﬀerent lengths, then R applies the recycling rule, If the list contains a NULL element, that element is skipped,


*** =instructions
- Use `cbind()`, `rbind()` in loops!

*** =hint
Follow the instruction and introduction. 

*** =pre_exercise_code
```{r}
```

*** =sample_code

```{r}
# create list of vectors
list_vectors <- lapply(1:5, rnorm, n=10)

# create matrix out of list


# see the dimension of matrix


# change the code to create function
do_call_rbind <- function(li_st) {
  while (length(li_st)) {
    odd_index <- seq(from=1, to=length(li_st), by=2)
    li_st <- lapply(odd_index, function(in_dex) {
      if (in_dex==length(li_st)) {
return(li_st[[in_dex]])
      }
      return(rbind(li_st[[in_dex]],
           li_st[[in_dex]]))
    })
  }
  li_st
}

# check if results are identical
identical(mat_rix, do_call_rbind(list_vectors))
```

*** =solution
```{r}
# create list of vectors
list_vectors <- lapply(1:5, rnorm, n=10)

# create matrix out of list
mat_rix <- do.call(rbind, list_vectors)

# see the dimension of matrix
dim(mat_rix)

# change the code to create function
do_call_rbind <- function(li_st) {
  while (length(li_st) > 1) {
    odd_index <- seq(from=1, to=length(li_st), by=2)
    li_st <- lapply(odd_index, function(in_dex) {
      if (in_dex==length(li_st)) {
return(li_st[[in_dex]])
      }
      return(rbind(li_st[[in_dex]],
           li_st[[in_dex+1]]))
    })
  }
  li_st[[1]]
}

# check if results are identical
identical(mat_rix, do_call_rbind(list_vectors))
```

*** =sct
```{r}
test_error()
success_msg("Yes! Now you know how to deal with a function!")
```
