--- 
title_meta  : Chapter 1
title       : R Environment
description : "This is the first chapter in the Introduction to R course. You will get to know the R environment. You will also learn basic manipulations on R objects, which is essential for R. Let's get started!"
attachments : 





--- type:NormalExercise lang:r xp:100 skills:1 key:5714863ba8
## Variables assignment

In the script window on the upper right you should type the *R* code required for the exercises. When you hit the *Submit Answer* button, your *R* code is executed, and you get a message indicating whether your *R* code is correct. The output of your submission is shown in the *R* console on the lower right.

You can also execute *R* commands in the console on the lower right. When you type in the console, your submission will not be checked for correctness! Try, for example, to type in `2 + 1` and hit Enter. *R* should return `[1] 3`.

Variables are created by an assignment operation, and they don’t have to be declared, The standard assignment operator in R is the arrow symbol `<-`, R interprets text in quotes (`””`) as character strings, Text that is not in quotes (`””`) is interpreted as a symbol or expression, Typing a symbol or expression evaluates it.

*** =instructions
In the script window on the upper right, assign number 2 and a string "Hello World!" to variables, and hit the *Submit Answer* button.

*** =hint
The variable assignments are already presented, just type the variable name and check results. Remember not to add quotes to them!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# assign number to a variable
myvariable1 <- 2

# assign a string to a variable
myvariable2 <- "Hello World!"

# now write R code of variable names to check if the assignments are successfull
```

*** =solution
```{r}
# assign number to a variable
myvariable1 <- 2

# assign a string to a variable
myvariable2 <- "Hello World!"

# now write R code to check if the assignments are successfull
myvariable1
myvariable2

```

*** =sct
```{r}
success_msg("Awesome! See how the console shows the result of the R code you submitted?")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:ab37530088
## Inserting comments into your code

*R* uses the hash `#` sign to mark text as comments.  
All text after the hash `#` sign is treated as a comment, and is not executed as code.  
For example, the text *Calculate 3 + 4* is a comment and is ignored during execution.

*** =instructions
Add a line with the comment mark *#* and  text *myvariable <- 1*, in the script window.

*** =hint
Simply add the line `# add comments to myvariable <- 1` above the *R* code that assigns values *myvariable1 <- 2*.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
myvariable <- 10

# assign number to a variable
myvariable1 <- 2

# assign a string to a variable
myvariable2 <- "Hello World!"

# add comments mark before assign 1 to 'myvariable1'


# now write R code to check if the assignments are successfull
```

*** =solution
```{r}
myvariable <- 10

# assign number to a variable
myvariable1 <- 2

# assign a string to a variable
myvariable2 <- "Hello World!"

# add comments
# myvariable <- 1

# now write R code to check if the assignments are successfull
myvariable
myvariable1
myvariable2
```

*** =sct
```{r}
success_msg("Awesome! See? That assignment in the comment does not take effect.")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:93f879e02e
## Explore sessions in R

R session has many default settings, which you can learn to configure or make use of.

The function getwd() returns a vector of length 1, with the ﬁrst element containing a string with the name of the current working directory (cwd).
The function setwd() accepts a character string as input (the name of the directory), and sets the working directory to that string. 
With Sys.time() you can get system time with time zone of the R session; or you can just get system date with Sys.Date().

*** =instructions
Use the function getwd(), Sys.time() and Sys.Date() to get strings of working directory, system time and system date.

*** =hint
Simply the function name, copy & paste.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# get string with the name of the current working directory


# get current system time with time zone


# get current date

```

*** =solution
```{r}
# get string with the name of the current working directory
getwd()

# get current system time with time zone
Sys.time()

# get current date
Sys.Date()
```


*** =sct
```{r}
success_msg("Great! Look at the results, play with them if you would like to")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:b1bb4402e2
## More about R sessions

The function sessionInfo() returns information about the current R session:
R version;
OS platform;
locale settings;
list of packages that are loaded and attached to the search path;
list of packages that are loaded, but not attached to the search path


*** =instructions
Use the function sessionInfo() to get detailed information about current R session.

*** =hint
Simply the function name, copy & paste.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# get information about the current R session

```

*** =solution
```{r}
# get information about the current R session
sessionInfo()
```


*** =sct
```{r}
success_msg("Correct! Remeber to see what's pop up in the console")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:2f547ffda2
## Environment Variables

R uses environment variables to store information about its environment, such as paths to directories containing ﬁles used by R (startup, history, OS).
For example the environment variables: 
R USER and HOME store the R user Home directory, 
R HOME stores the root directory of the R installation, 
The functions Sys.getenv() and Sys.setenv() display and set the values environment variables, 
Sys.getenv("env var") displays the environment variable "env var", 
Sys.setenv("env var=value") sets the environment variable "env var" equal to "value",


*** =instructions
Use `Sys.getenv()` to display values of environment variables.
/nChange parameters as instructed and see the results.

*** =hint
Simply copy the function name and change parameters inside.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# get the subset from 5th to 7th element for list of environment variable by subsetting
Sys.getenv()[5:7]

# get user HOME directory 


# get R_HOME directory 


# get "etc" sub-directory of R_HOME

```

*** =solution
```{r}
# get the subset from 5th to 7th element for list of environment variable by subsetting
Sys.getenv()[5:7]

# get user HOME directory 
Sys.getenv("Home")

# get R_HOME directory 
Sys.getenv("R_home")

# get "etc" sub-directory of R_HOME
Sys.getenv("env_var")
```


*** =sct
```{r}
success_msg("Correct! Remeber to see what's pop up in the console")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:4d04bf29c2
## Global Options Settings

R uses a list of global options which aﬀect how R computes and displays results./n
The function options() either sets or displays the values of global options:/n 
  options("globop") displays the current value of option "globop";/n
  getOption("globop") displays the current value of option "globop";/n
  options(globop=value) sets the option "globop" equal to "value";


*** =instructions
Change parameters as instructed and see the results.

*** =hint
Remeber to capitalize TRUE and don't use commas for the last code!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# get number of digits printed for numeric values
options("digits")

# get the current value of "digits"


# set the digits option to 3, type 0.12345 and see results

```

*** =solution
```{r}
# get number of digits printed for numeric values
getOption("digits")

# get the current value of "digits"
options("digits")

# set the digits option to 3, type 0.12345 and see results
options(digits = 3)
0.12345
```


*** =sct
```{r}
success_msg("OK, now see what's pop up in the console")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:58ca8cb732
## The R Workspace

The workspace is the current R working environment, which includes all user-deﬁned objects and the command history./n
The function ls() returns names of objects in the R workspace./n
The function rm() removes objects from the R workspace./n
The workspace can be saved into and loaded back from an *.RData ﬁle (binary ﬁle format)./n
The function save.image() saves the whole workspace. /n
The function save() saves just the selected objects./n
The function load() reads data from *.RData ﬁles, and invisibly returns a vector of names of objects created in the workspace./n

When you quit R you’ll be prompted ”Save workspace image?” If you answer YES then the workspace will be saved into the .RData ﬁle in the cwd./n
When you start R again, the workspace will be automatically loaded from the existing .RData ﬁle./n
/n
The function history() displays recent commands, You can also save and load the command history from a ﬁle./n

*** =instructions
Use the function ls() and rm() to see all objects in the environment and remove some of them.

*** =hint
Simply the function name, copy & paste as instructed by comments.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
variable_1 <- 5
variable_2 <- "sand"

# see what variable is currently in the workspace


# remove all elements in the workspace
rm(list = ls())

# check what is left in the workspace

```

*** =solution
```{r}
variable_1 <- 5
variable_2 <- "sand"
# see what variable is currently in the workspace
ls()

# remove all elements in the workspace
rm(list = ls())

# check what is left in the workspace
ls()
```


*** =sct
```{r}
success_msg("Great! Look at the results, play with them if you would like to")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:39d8f62180
## Data Objects in R

All data objects in R are vectors, or consist of vectors.
Single numbers and character strings are vectors of length "1".
Atomic vectors are homogeneous objects whose elements are all of the same mode (type).
Lists and data frames are recursive (heterogeneous) objects, whose elements can be vectors of diﬀerent mode.
The functions is.atomic() and is.recursive() return logical values depending on whether their arguments are atomic or recursive.

*** =instructions
Use function is.atomic() and is.recursive() to check whether the object is atomic or recursive.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# define objects of single number, vector, list and dataframe
v1 <- 1
v2 <- 1:10
v3 <- list(aa=c('a', 'b'), bb=1:5)
v4 <- data.frame(aa=c('a', 'b'), bb=1:2)

# find out if v1, v2 is atomic or recursive
is.atomic(v1)
is.recursive(v1)

# find out if v2 (vector) is atomic or recursive


# use similar process to find out if v3 (list) is atomic or recursive


# do the same to v4 (dataframe)

```

*** =solution
```{r}
# define objects of single number, vector, list and dataframe
v1 <- 1
v2 <- 1:10
v3 <- list(aa=c('a', 'b'), bb=1:5)
v4 <- data.frame(aa=c('a', 'b'), bb=1:2)

# find out if v1, v2 is atomic or recursive
is.atomic(v1)
is.recursive(v1)

# find out if v2 (vector) is atomic or recursive
is.atomic(v2)
is.recursive(v2)

# use similar process to find out if v3 (list) is atomic or recursive
is.atomic(v3)
is.recursive(v3)

# do the same to v4 (dataframe)
is.atomic(v4)
is.recursive(v4)
```


*** =sct
```{r}
success_msg("You can scroll down and console and see the output")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:a5293e3bde
## Type, Mode, and Class of Objects

The type, mode, and class are character strings representing various object properties.
The type of an atomic object represents how it’s stored in memory (logical, character, integer, double, etc.)
The mode of an atomic object is the kind of data it represents (logical, character, numeric, etc.)
The mode of an object often coincides with its type (except for integer and double types)
Recursive objects (lists, data frames) have both type and mode equal to the recursive type (list)
The class of an object is given by either an explicit class attribute, or is derived from the object mode and its dim attribute (implicit class)
The function class() returns the explicit or implicit class of an object.
The object class is used for method dispatching in the S3 object-oriented programming system in R

*** =instructions
Use function typeof(), mode() and class() to explore object type, mode and class.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# define objects of vector, matrix, list and dataframe
v1 <- 1:10
v2 <- matrix(1:10, 2, 5)
v3 <- list(aa=c('a', 'b'), bb=1:5)
v4 <- data.frame(aa=c('a', 'b'), bb=1:2)

# find out the type, mode and class of vector and matrix
c(typeof(v1), mode(v1), class(v1))
c(typeof(v2), mode(v2), class(v2))

# use the same code to check list and dataframe

```

*** =solution
```{r}
# define objects of vector, matrix, list and dataframe
v1 <- 1:10
v2 <- matrix(1:10, 2, 5)
v3 <- list(aa=c('a', 'b'), bb=1:5)
v4 <- data.frame(aa=c('a', 'b'), bb=1:2)

# find out the type, mode and class of vector and matrix
c(typeof(v1), mode(v1), class(v1))
c(typeof(v2), mode(v2), class(v2))

# now use the same code to check list and dataframe
c(typeof(v3), mode(v3), class(v3))
c(typeof(v4), mode(v4), class(v4))
```


*** =sct
```{r}
success_msg("Good Job!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:85b773db52
## R Object Attributes

R objects can have diﬀerent attributes, such as: names, dimnames, dimensions, class, etc.
The attributes of an object is a named list of symbol=value pairs.
The function attributes() returns the attributes of an object.
The attributes of an R object can be modiﬁed using the "attributes()" assignment.

The function structure() adds attributes (speciﬁed as symbol=value pairs) to an object, and returns it.
A vector that is assigned an attribute other than names is not treated as a vector.
The function is.vector() returns TRUE if its argument is a vector, and returns FALSE otherwise.


*** =instructions
Use function attributes(), structure() and is.vector() to explore object attributes.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# check the attributes of a simple vector 1:10


my_var <- c(pi=pi, euler=exp(1), gamma=-digamma(1))
# check the attributes of names vector `my_var`


# modify the attributes of an object, convert the class to "Date"


# print the object and see what's changed


# use structure to create a vector of Date from 1:10


my_list <- list(1, 2)
# now use attributes assignment method to change the attributes of my_list to a vector of Dates


# print out your result


# check if it is a vector

```

*** =solution
```{r}
# check the attributes of a simple vector 1:10
attributes(1:10)

my_var <- c(pi=pi, euler=exp(1), gamma=-digamma(1))
# check the attributes of names vector `my_var`
attributes(my_var)

# modify the attributes of an object, convert the class to "Date"
attributes(my_var) <- list(class = "Date")

# print the object and see what's changed
my_var

# use structure to create a vector of Date from 1:10
structure(1:10, class = "Date")

my_list <- list(1, 2)
# now use attributes assignment method to change the attributes of my_list to a vector of Dates
attributes(my_list) <- list(class = "Date")

# print out your result
my_list

# check if it is a vector
is.vector(my_list)
```


*** =sct
```{r}
test_error()
msg <- "Do not remove or change the definition of the variable `my_list`."
test_object("my_list", undefined_msg = msg, incorrect_msg = msg)
test_function("is.vector", "x",
              not_called_msg = "Make sure to call the function [`is.vector()`] to see if the object is a vector.",
              incorrect_msg = "Have you passed the correct variable to the function [`is.vector()`]?")
success_msg("Good Job!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:792449dde3
## Implicit Class of Objects

We will revisit the class of objects in this exercise. This time there are more details:

If an object has no explicit class, then its implicit class is derived from its mode and dim attribute (except for integer vectors which have the implicit class "integer" derived from their type).

If an atomic object has a dim attribute, then its implicit class is matrix or array.
Data frames have no explicit dim attribute, but dim() returns a value, so they have an implicit dim attribute.


*** =instructions
Now that a dataframe is created. Use function typeof(), mode(), class() and attributes() to explore results.

*** =hint
Check if object names and function names are correctly typed. Remember to use all of the functions in the instructions.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# integer implicit class derived from type
my_var_1 <- vector(mode="integer", length=10)

# check the type, mode and class of my_var1


# numeric implicit class derived from mode
my_var_2 <- vector(mode="numeric", length=10)

# check the type, mode and class of my_var2


# add c(5,2) to dim attribute of `my_var_1` changes implicit class to matrix


# check the type, mode and class of my_var1


# define a new dataframe
my_var <- data.frame(aa=c('a', 'b'), bb=1:2)

# now explore the type, class, mode of a data frame


# check the attributes of my_var

```

*** =solution
```{r}
# integer implicit class derived from type
my_var_1 <- vector(mode="integer", length=10)

# check the type, mode and class of my_var1
c(typeof(my_var_1), mode(my_var_1), class(my_var_1))

# numeric implicit class derived from mode
my_var_2 <- vector(mode="numeric", length=10)

# check the type, mode and class of my_var2
c(typeof(my_var_2), mode(my_var_2), class(my_var_2))

# add c(5,2) to dim attribute of `my_var_1` changes implicit class to matrix
dim(my_var_1) <- c(5, 2)

# check the type, mode and class of my_var1
c(typeof(my_var_1), mode(my_var_1), class(my_var_1))

# define a new dataframe
my_var <- data.frame(aa=c('a', 'b'), bb=1:2)

# now explore the type, class, mode of a data frame
c(typeof(my_var), mode(my_var), class(my_var))

# check the attributes of my_var
attributes(my_var)
```


*** =sct
```{r}
test_error()
msg <- "Do not remove or change the definition of the variable `my_var`."
test_object("my_var", undefined_msg = msg, incorrect_msg = msg)
#test_function("typeof", "x",
#              not_called_msg = "Make sure to call the function [`typeof()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`typeof()`]?")
#test_function("mode", "x",
#              not_called_msg = "Make sure to call the function [`mode()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`mode()`]?")
#test_function("class", "x",
#              not_called_msg = "Make sure to call the function [`class()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`class()`]?")
#test_function("attributes", "x",
#              not_called_msg = "Make sure to call the function [`attributes()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`attributes()`]?")
success_msg("Good Job!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:914dcdc4c2
## Object Coercion

Coercion means changing the type, mode, or class of an object, often without changing the underlying data. 
Changing the mode of an object can change its class as well, but not always.
Objects can be explicitly coerced using the "as.*" coercion functions.
Most coercion functions strip the attributes from the object.
Implicit coercion occurs when objects with diﬀerent modes are combined into a vector, forcing the elements to have the same mode.
Implicit coercion can cause bugs that are diﬃcult to trace.
The rule is that coercion is into larger types (numeric objects are coerced into character strings).

Notice: Coercion can introduce bad data, such as NA values.


*** =instructions
Use coercison to convert object types with functions of `as.*`.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# create a numeric vector
my_var_1 <- vector(mode="integer", length=10)

# check the type, mode and class of my_var_1


# coerce the vector into character


# check the type, mode and class of my_var_1 after the change


# create a mixed vector
my_var <- c(2, "a", 3.2)

# check the type, mode and class of my_var


# now coerce the vector into integer


# check the type, mode and class of my_var

```

*** =solution
```{r}
# create a numeric vector
my_var_1 <- vector(mode="integer", length=10)

# check the type, mode and class of my_var_1
c(typeof(my_var_1), mode(my_var_1), class(my_var_1))

# coerce the vector into character
my_var_1 <- as.character(my_var_1)

# check the type, mode and class of my_var_1 after the change
c(typeof(my_var_1), mode(my_var_1), class(my_var_1))

# create a mixed vector
my_var <- c(2, "a", 3.2)

# check the type, mode and class of my_var
c(typeof(my_var), mode(my_var), class(my_var))

# now coerce the vector into integer
my_var <- as.integer(my_var)

# check the type, mode and class of my_var
c(typeof(my_var), mode(my_var), class(my_var))
```


*** =sct
```{r}
test_error()
msg <- "Do not remove or change the definition of the variable `my_var`."
test_object("my_var", undefined_msg = msg, incorrect_msg = msg)
#test_function("typeof", "x",
#              not_called_msg = "Make sure to call the function [`typeof()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`typeof()`]?")
#test_function("mode", "x",
#              not_called_msg = "Make sure to call the function [`mode()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`mode()`]?")
#test_function("class", "x",
#              not_called_msg = "Make sure to call the function [`class()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`class()`]?")
#test_function("attributes", "x",
#              not_called_msg = "Make sure to call the function [`attributes()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`attributes()`]?")
success_msg("Good Job!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:b1b1305012
## Basic R Objects

The quotation marks "" (or ’’) around a character string tell R that it’s a string, not a variable name.
Vectors are the basic building blocks of R objects, There are no scalars in R, and single values are stored as vectors of length "1".
A character string is also a vector with a single element, with the ﬁrst element of the vector containing the string of text.
The colon binary operator ’:’ produces a vector.
The function c() combines objects into a vector.
The "[1]" symbol means the return value is a vector.
The function is.vector() returns TRUE if its argument is a vector, and returns FALSE otherwise,


*** =instructions
Try to explore different kinds of vector and make use of vector's basic operations.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# create a character vector
vec_char <- "some string"

# explore the vector with basic operations of vector
is.vector(vec_char)
vec_char[1]

# create a numeric vector
vec_num <- 1:10

# print and explore the vector with basic operations of vector
vec_num
is.vector(vec_num)
vec_num[2]

#now use c() to combine vec_char and vec_num and create vec_c


#check if vec_c is a vector and print the second element of vec_c
```

*** =solution
```{r}
# create a character vector
vec_char <- "some string"

# explore the vector with basic operations of vector
is.vector(vec_char)
vec_char[1]

# create a numeric vector
vec_num <- 1:10

# print and explore the vector with basic operations of vector
vec_num
is.vector(vec_num)
vec_num[2]

#now use c() to combine vec_char and vec_num and create vec_c
vec_c <- c(vec_char, vec_num)

#check if vec_c is a vector and print the second element of vec_c
is.vector(vec_c)
vec_c[2]
#
```


*** =sct
```{r}
test_error()
msg <- "Do not remove or change the definition of the variable `vec_num`."
test_object("vec_num", undefined_msg = msg, incorrect_msg = msg)
success_msg("Great!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:05697caf6a
## Character Strings

Character strings are sequences of characters (and vectors of length one), The function nchar() returns the length of a string.

Special characters in strings: 
"\t" for TAB, "\n" for new-line, 
”\\” for a (single) backslash character

The function cat() concatenates strings and echos them to console, without returning any values
The function cat() is useful in user-deﬁned functions.


*** =instructions
Explore characters and strings with functions and basic features.

*** =hint
Check if object names and function names are correctly typed. Don't miss any variables or functions!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
#create a character string
str_var <- "Some string"

#print the entire vector and its first element
str_var
str_var[1]

#try to print its second element


#see the length of vector and string
length(str_var)
nchar(str_var)

#concatenate and echo to console
cat("Hello", "World!")

#use \n to start a new line when concatenating

```

*** =solution
```{r}
#create a character string
str_var <- "Some string"

#print the entire vector and its first element
str_var
str_var[1]

#try to print its second element
str_var[2]

#see the length of vector and string
length(str_var)
nchar(str_var)

#concatenate and echo to console
cat("Hello", "World!")

#use \n to start a new line when concatenating
cat("Enter\nnewline")
```


*** =sct
```{r}
test_error()
msg <- "Do not remove or change the definition of the variable `str_var`."
test_object("str_var", undefined_msg = msg, incorrect_msg = msg)
success_msg("Excellent!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:b99299073f
## Manipulating Strings

The function paste() concatenates its arguments into a string, coerces them to characters if needed, and returns the string.
If a vector or list is passed to paste(), together with a collapse string, then paste() concatenates the elements into a string, separated by the collapse string.
The function strsplit() splits the elements of a character vector.
Splitting on the "." character requires surrounding it with brackets: "[.]", or using argument fixed=TRUE.
The function substring() extracts or replaces substrings in a character string, The recycling rule extends the length to match the longest object.

*** =instructions
Use functions to manipulate strings.

*** =hint
Check if object names and function names are correctly typed. Don't miss the square bracket!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
#create two character string and combine
str_var1 <- "Hello"
str_var2 <- "World!"
paste(str_var1, str_var2, sep=' ')

#create a character string and a numeric vector
str_var3 <- "a"
num_var <- 1:4

#try to combine them together with "," and see results


#create a new string
hello <- "Hello.World"

#split the string into seperate strings using strsplit() on "[.]"


#use substring to get from the third to the sixth elements

```

*** =solution
```{r}
#create two character string and combine
str_var1 <- "Hello"
str_var2 <- "World!"
paste(str_var1, str_var2, sep=' ')

#create a character string and a numeric vector
str_var3 <- "a"
num_var <- 1:4

#try to combine them together with "," and see results
paste(str_var3, num_var, sep = ",")

#create a new string
hello <- "Hello.World"

#split the string into seperate strings using strsplit() on "[.]"
strsplit(hello, split = "[.]")

#use substring to get from the third to the sixth elements
substring(hello, 3, 6)
```


*** =sct
```{r}
test_error()
success_msg("Excellent!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:ab03e239aa
## Regular Expressions in R

R has Regex functions for pattern matching and replacement.
The function gsub() replaces all matches of a pattern in a string.
The function grep() searches for matches of a pattern in a string.
The function glob2rx() converts globbing wildcard patterns into regular expressions.

*** =instructions
Use functions to manipulate strings.

*** =hint
Check if object names and function names are correctly typed. Don't forget to use "i+" instead of "i"

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
#replace "is" with "XX"
my_string <- c("is", "XX", "is this gratis?")
gsub(my_string)

#get indexes of elements in a vector that contains certain string
grep("b+", c("abc", "xyz", "cba d", "bbb"))

#get values among elements in a vector that contains certain string
grep("b+", c("abc", "xyz", "cba d", "bbb"), value=TRUE)

# convert globs into regex (regular expression)
glob2rx("abc.*")

#now try to get the index of elements that contain character "i" in my_string


#convert my_string to regex

```

*** =solution
```{r}
#replace "is" with "XX"
my_string <- c("is", "XX", "is this gratis?")
gsub("is", "XX", my_string)

#get indexes of elements in a vector that contains certain string
grep("b+", c("abc", "xyz", "cba d", "bbb"))

#get values among elements in a vector that contains certain string
grep("b+", c("abc", "xyz", "cba d", "bbb"), value=TRUE)

# convert globs into regex (regular expression)
glob2rx("abc.*")

#now try to get the index of elements that contain character "i+" in my_string
grep("i+", my_string)

#convert my_string to regex
glob2rx(my_string)
```


*** =sct
```{r}
test_error()
success_msg("Remember to look at what's popped up in the console")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:540b8c57a5
## Vectors

Vectors are the basic building blocks of R objects.

For a brief review, we already know that:
There are no scalars in R, and single values are stored as vectors of length "1"
The function c() combines values into a vector
The function is.vector() returns TRUE if its argument is a vector, and returns FALSE otherwise

Beside, we need to knwo that the object letters is a constant and a vector.

To create a vector, R provides several way to do it:
1) The colon operator (":") provides a simple way of creating a numeric vector
2) The function vector() returns a vector of the speciﬁed mode
3) The functions seq(), seq_len(), and seq_along() return a sequence (vector) of numbers
4) The function rep() replicates an object multiple times
5) The functions numeric() and character() return empty (zero length) vectors of the speciﬁed type (not to be confused with a NULL object)
6) Zero length vectors are not NULL objects,


*** =instructions
Use mentioned functions to create vector.

*** =hint
Check if object names and function names are correctly typed. Use "by=" instead of "len=".

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
#create vectors using ":", vector() and rep()
1:10
vector(mode = "numeric", length = 10)
rep(10, times = 5)

#create vectors using seq
seq(10)
seq(along=(-5:5))
seq_along(c("a", "b", "c"))
seq(from=0, to=1, len=11)

#now try to use seq() to create a vector from 0 to 1, each step is 0.1, using "by="


#use character() to create empty vector of characters
character(5)

#now try to use numeric() to create empty vector of numerics of length 0

```

*** =solution
```{r}
#create vectors using ":", vector() and rep()
1:10
vector(mode = "numeric", length = 10)
rep(10, times = 5)

#create vectors using seq
seq(10)
seq(along=(-5:5))
seq_along(c("a", "b", "c"))
seq(from=0, to=1, len=11)

#now try to use seq() to create a vector from 0 to 1, each step is 0.1, using "by="
seq(from = 0, to = 1, by = 0.1)

#use character() to create empty vector of characters
character(5)

#now try to use numeric() to create empty vector of numerics of length 0
numeric(0)
```


*** =sct
```{r}
test_error()
success_msg("See? Empty vector with zero length is not equal to NULL!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:5a2dc65a8a
## Arithmetic and Logical Operations on Vectors

Vectors can be multiplied and squared element by element, as if they were single elements.

When vectors are manipulated as if they were single elements, then R automatically performs a loop over the vector elements, and applies the operation element-wise. This is a very powerful feature of R called vectorized arithmetic. Vectorized arithmetic avoids writing loops and simpliﬁes notation.

Vectors can be combined together and appended.

*** =instructions
Use mentioned functions to conduct vector arithmetic and logical operation.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
#multiply a vector
2*4:8
2*(4:8)

#divide a vector
4:8/2
(0:10)/10

#create vector
vec_tor <- c(8, 6, 5, 7)
vec_tor

#boolean vector TRUE if element is equal to second one, otherwise FALSE
vec_tor == vec_tor[2]

#try to create boolean vector TRUE for elements greater than six, otherwise FALSE


#square all elements
vec_tor^2

#now try to multiply all elements by 2


#combine two vectors
c(11, 5:10)

#now try to append number to vector

```

*** =solution
```{r}
#multiply a vector
2*4:8
2*(4:8)

#divide a vector
4:8/2
(0:10)/10

#create vector
vec_tor <- c(8, 6, 5, 7)
vec_tor

#boolean vector TRUE if element is equal to second one
vec_tor == vec_tor[2]

#try to create boolean vector TRUE for elements greater than six
vec_tor > 6

#square all elements
vec_tor^2

#now try to multiply all elements by 2
2*vec_tor

#combine two vectors
c(11, 5:10)

#now try to append number to vector
c(vec_tor, 2.0)
```


*** =sct
```{r}
test_error()
success_msg("Beautiful! Appreciate results in the console!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:2635635720
## Naming and Manipulating Vectors

Vector elements can be assigned names using a list of symbol-value pairs.
The function names() returns the names attribute of an object.
The names attribute of a vector can be modiﬁed by assigning to the names() function (”names() <-” assignment).
The function unname() removes the names attribute.
The function structure() adds attributes (speciﬁed as symbol=value pairs) to an object, and returns it,

*** =instructions
Use names() to get names and assign names. Also learn to use structure() to assign names to a vector.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# create named vector
vec_tor <- c(pi_const=pi, euler=exp(1), gamma=-digamma(1))

# get names of elements
names(vec_tor)

# remove names attribute
unname(vec_tor)

# combine two vectors of letters
c('a', letters[5:10])

# create named vector using structure
structure(sample(1:5), names=paste0("el", 1:5))

# now try to assign 1:3 to names of vec_tor


# print out vec_tor


# use structure to create a vector of 1:3, naming it with vec_tor

```

*** =solution
```{r}
# create named vector
vec_tor <- c(pi_const=pi, euler=exp(1), gamma=-digamma(1))

# get names of elements
names(vec_tor)

# remove names attribute
unname(vec_tor)

# combine two vectors of letters
c('a', letters[5:10])

# create named vector using structure
structure(sample(1:5), names=paste0("el", 1:5))

# now try to assign 1:3 to names of vec_tor
names(vec_tor) <- 1:3

# print out vec_tor
vec_tor

# use structure to create a vector of 1:3, naming it by vec_tor
structure(1:3, names = vec_tor)
```


*** =sct
```{r}
test_error()
success_msg("Now you can name a vector with names() and structure() methods!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:e2b1d979f4
## Subsetting Vectors

Vector elements can be subset (indexed, dereferenced) using the "[]" operator.

Vectors can be subset using vectors of:
positive integers, 
negative integers, 
characters (names),
boolean vectors. 

Negative integers remove the vector elements.
Subsetting with zero returns a zero-length vector.
A named vector can be subset using element names.

*** =instructions
Vec_tor has been recreated in this exercise. 
Use "[]" combined with integers, characters and boolean vectors to subset a vector.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
vec_tor <- c(pi_const=pi, euler=exp(1), gamma=-digamma(1))
```

*** =sample_code
```{r}
# extract second element
vec_tor[2]

# extract all elements, except the second element
vec_tor[-2]

# extract zero elements - returns zero-length vector
vec_tor[0]

# extract second and third elements
vec_tor[c(FALSE, TRUE, TRUE)]

# extract a element using their names
vec_tor["eulery"]

# extract multiple elements using a vector of their names
vec_tor[c("pie", "gammy")]

# now create a boolean vector comparing vec_tor to 2, name it as "bool"


# subset vec_tor with bool to see which element is bigger than 2

```

*** =solution
```{r}
# extract second element
vec_tor[2]

# extract all elements, except the second element
vec_tor[-2]

# extract zero elements - returns zero-length vector
vec_tor[0]

# extract second and third elements
vec_tor[c(FALSE, TRUE, TRUE)]

# extract a element using their names
vec_tor["eulery"]

# extract multiple elements using a vector of their names
vec_tor[c("pie", "gammy")]

# now create a boolean vector comparing vec_tor to 2, name it as "bool"
bool <- vec_tor > 2

# subset vec_tor with bool to see which element is bigger than 2
vec_tor[bool]
```


*** =sct
```{r}
test_error()
success_msg("The subsetting provides support for filtering")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:47e9ca9b81
## Filtering Vectors

Filtering means extracting elements from a vector that satisfy a logical condition.
When logical comparison operators are applied to vectors, they produce boolean vectors.
Boolean vectors can then be applied to subset the original vectors, to extract their elements.
The function which() returns the indices of the TRUE elements of a boolean vector or array.


*** =instructions
Vec_tor has been recreated in this exercise. 
Use "[]"  or "which()" to filter vector elements.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
vec_tor <- runif(5)
```

*** =sample_code
```{r}
# first, print the vec_tor
vec_tor

# compare vec_tor to 0.5, in this way you actually created a boolean vector
vec_tor > 0.5

# boolean vector of elements equal to the second one
vec_tor == vec_tor[2]

# extract all elements equal to the second one
vec_tor[vec_tor == vec_tor[2]]

# now extract all elements greater than 0.5


# use which() and returns the index of filtered element greater than 0.5
which(vec_tor > 0.5)

# use which() and "[]" to extract elements greater than 0.5

```

*** =solution
```{r}
# first, print the vec_tor
vec_tor

# compare vec_tor to 0.5, in this way you actually created a boolean vector
vec_tor > 0.5

# boolean vector of elements equal to the second one
vec_tor == vec_tor[2]

# extract all elements equal to the second one
vec_tor[vec_tor == vec_tor[2]]

# now extract all elements greater than 0.5
vec_tor[vec_tor > 0.5]

# use which() and returns the index of filtered element greater than 0.5
which(vec_tor > 0.5)

# use which() and "[]" to extract elements greater than 0.5
vec_tor[which(vec_tor > 0.5)]
```


*** =sct
```{r}
test_error()
success_msg("Now you can filter a vector with '[]' and which()!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:ab3a8365ff
## Factors

Factors are similar to vectors, but their elements can only take values from a set of levels.
Factors are designed for categorical data which can only take certain values.

The function factor() converts a vector into a factor, Factors have two attributes: class (equal to "factor") and levels (the allowed values).
Although factors aren’t vectors, the data underlying a factor is an integer vector, called an encoding vector.
The function as.numeric() extracts the encoding vector (indices) of a factor.
The function as.vector() coerces a factor to a character vector.

*** =instructions
Create factors with factor(). 
See attributes of factor using attributes() and levels().
Coerce factor using as.numeric() and as.vector().

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
fac_tor <- factor(c('b', 'c', 'd', 'a', 'c', 'b'))
```

*** =sample_code
```{r}
# have a look at the factor vector
fac_tor

# factor can be subsetted the same way using "[]"
fac_tor[3]

# get factor attributes
attributes(fac_tor)

# get factor levels
levels(fac_tor)

# get encoding vector
as.numeric(fac_tor)

# coerce vector of 1:5 to factor, and then coerce it back using as.vector()

```

*** =solution
```{r}
# have a look at the factor vector
fac_tor

# factor can be subsetted the same way using "[]"
fac_tor[3]

# get factor attributes
attributes(fac_tor)

# get factor levels
levels(fac_tor)

# get encoding vector
as.numeric(fac_tor)

# coerce vector of 1:5 to factor, and then coerce it back using as.vector()
as.vector(factor(1:5))
```


*** =sct
```{r}
test_error()
success_msg("See! The numeric vector turned to be a character vector!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:771f2e9f50
## Tables of Categorical Data

The function table() calculates the frequency distribution of categorical data.
A contingency table is a matrix that contains the frequency distribution of variables (factors) contained in a set of data, sapply() applies a function to a vector or a list of objects and returns a vector or a list,


*** =instructions
Use functions and variables as insturcted in the comment.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
fac_tor <- factor(c('b', 'c', 'd', 'a', 'c', 'b'))
```

*** =sample_code
```{r}
# have a look at the factor vector
fac_tor

# get encoding vector
as.numeric(fac_tor)

# get unique values
unique(fac_tor)

# get contingency table using sapply 
sapply(levels(fac_tor), 
 function(le_vel) {
   sum(fac_tor==le_vel)
 })

# get contingency (frequency) table using table() on fac_tor

```

*** =solution
```{r}
# have a look at the factor vector
fac_tor

# get encoding vector
as.numeric(fac_tor)

# get unique values
unique(fac_tor)

# get contingency table using sapply 
sapply(levels(fac_tor), 
 function(le_vel) {
   sum(fac_tor==le_vel)
 })

# get contingency (frequency) table using table() on fac_tor
table(fac_tor)
```


*** =sct
```{r}
test_error()
success_msg("Compared the results you get with sapply() and table()")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:748763cc09
## Classifying Continuous Numeric Data Into Categories

Numeric data that represents a magnitude, intensity, or score can be classiﬁed into categorical data, given a vector of breakpoints.

The breakpoints create intervals that correspond to diﬀerent categories.

The categories combine elements that have a similar numeric magnitude, findInterval() returns the indices of the intervals speciﬁed by "vec" that contain the elements of "x".
If there’s an exact match, then findInterval() returns the same index as function match().
If there’s no exact match, then findInterval() ﬁnds the element of "vec" that is closest to, but not greater than, the element of "x".
If all the elements of "vec" are greater than the element of "x", then findInterval() returns zero,


*** =instructions
Use functions and variables as insturcted in the comment.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# none
```

*** =sample_code
```{r}
# see how we can tweak the outcome findInterval() function with five parameters
str(findInterval)

# get index of the element of "vec" that matches 5
findInterval(x=5, vec=c(3, 5, 7))
match(5, c(3, 5, 7))

# no exact match
findInterval(x=6, vec=c(3, 5, 7))
match(6, c(3, 5, 7))

# indices of "vec" that match elements of "x"
findInterval(x=1:8, vec=c(3, 5, 7))

# return only indices of inside intervals


# make rightmost interval inclusive

```

*** =solution
```{r}
# see how we can tweak the outcome findInterval() function with five parameters
str(findInterval)

# get index of the element of "vec" that matches 5
findInterval(x=5, vec=c(3, 5, 7))
match(5, c(3, 5, 7))

# no exact match
findInterval(x=6, vec=c(3, 5, 7))
match(6, c(3, 5, 7))

# indices of "vec" that match elements of "x"
findInterval(x=1:8, vec=c(3, 5, 7))

# return only indices of inside intervals
findInterval(x=1:8, vec=c(3, 5, 7), all.inside=TRUE)

# make rightmost interval inclusive
findInterval(x=1:8, vec=c(3, 5, 7), rightmost.closed=TRUE)
```


*** =sct
```{r}
test_error()
success_msg("Congrats! This is the very first step in dividing continuous values")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:91aa35a650
## Classifying Continuous Numeric Data Into Categories, Continued

Temperature can be categorized into ”cold”, ”warm”, ”hot”, etc. 
A named numeric vector of breakpoints can be used to convert a temperature into one of the categories.
Breakpoints correspond to categories of the data.
The ﬁrst breakpoint should correspond to the lowest category, and should have a value less than any of the data.


*** =instructions
Use functions and variables as insturcted in the comment.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# none
```

*** =sample_code
```{r}
# creat named numeric vector of breakpoints
brea_ks <- c("freezing"=0, "very_cold"=30,
       "cold"=50, "pleasant"=60,
       "warm"=80, "hot"=90)

# print out the breakpoints vector
brea_ks

# create a vector of temperatures
tempe_ratures <- runif(10, min=10, max=100)

# name values according to breakpoints
feels_like <- names(
  brea_ks[findInterval(x=tempe_ratures,
                 vec=brea_ks)])

names(tempe_ratures) <- feels_like

# print the categorized vector

```

*** =solution
```{r}
# creat named numeric vector of breakpoints
brea_ks <- c("freezing"=0, "very_cold"=30,
       "cold"=50, "pleasant"=60,
       "warm"=80, "hot"=90)

# print out the breakpoints vector
brea_ks

# create a vector of temperatures
tempe_ratures <- runif(10, min=10, max=100)

# name values according to breakpoints
feels_like <- names(
  brea_ks[findInterval(x=tempe_ratures,
                 vec=brea_ks)])

names(tempe_ratures) <- feels_like

# print the categorized vector
tempe_ratures
```


*** =sct
```{r}
test_error()
success_msg("Congrats! This is the your first value categorization results!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:ca09ff3670
## Converting Numeric Data Into Factors Using cut()

The function cut() converts a numeric vector into a vector of factors, representing the intervals to which the numeric values belong, cut() divides the range of values into intervals, based on a vector of breaks, cut() then assigns factors to the numeric values, representing the intervals to which the numeric values belong. 
The argument "breaks" is a numeric vector of break points that divide the range of values into intervals.
The argument "labels" is a vector of labels for the intervals.
The argument "right" is a boolean indicating if the intervals should be closed on the right (and open on the left), or vice versa, cut() can produce the same classiﬁcation as findInterval(), but findInterval() is faster than cut(), because it’s a compiled function,

The library microbenchmark is already loaded to the environment

*** =instructions
Use functions and variables as insturcted in the comment.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
library(microbenchmark)
```

*** =sample_code
```{r}
# create a vector of random value
foo <- sample(0:6) + 0.1

# print the vector
foo

# use cut to vector
cut(x=foo, breaks=c(2, 4, 6, 8))
rbind(foo, cut(x=foo, breaks=c(2, 4, 6, 8)))

# use findInterval() to cut the vector and compare the results


# findInterval() is a compiled function, so it is faster than cut()
vec_tor <- rnorm(1000)
summary(microbenchmark(
  find_interval=
    findInterval(x=vec_tor, vec=c(3, 5, 7)),
  cuut=
    cut(x=vec_tor, breaks=c(3, 5, 7)),
  times=10))[, c(1, 4, 5)]
```

*** =solution
```{r}
# create a vector of random value
foo <- sample(0:6) + 0.1

# print the vector
foo

# use cut to vector
cut(x=foo, breaks=c(2, 4, 6, 8))
rbind(foo, cut(x=foo, breaks=c(2, 4, 6, 8)))

# use findInterval() to cut the vector and compare the results
findInterval(x=1:8, vec=c(2, 4, 6, 8))

# findInterval() is a compiled function, so it is faster than cut()
vec_tor <- rnorm(1000)
summary(microbenchmark(
  find_interval=
    findInterval(x=vec_tor, vec=c(3, 5, 7)),
  cuut=
    cut(x=vec_tor, breaks=c(3, 5, 7)),
  times=10))[, c(1, 4, 5)]
```


*** =sct
```{r}
test_error()
success_msg("Althogh slower and a little bit different, cut() has similar functionality as findInternval()")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:cd32ce976a
## Matrices

The function matrix() creates a matrix from a vector, and the matrix dimensions.
By default matrix() creates matrices column-wise, unless the argument byrow=TRUE is used.
The elements of matrices can be subset (dereferenced) using the "[]" operator.
The functions nrow() and ncol() return the number of rows and columns of a matrix.
The functions NROW() and NCOL() also return the number of rows or columns of a matrix, but they can also be applied to vectors, and treat vectors as single column matrices.

*** =instructions
Use matrix() to construct matrices, and nrow(), ncol() and NROW(), NCOL().

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
vec_tor <- runif(5)
```

*** =sample_code
```{r}
# create a matrix
mat_rix <- matrix(5:10, nrow=2, ncol=3)

# by default matrices are constructed column-wise
mat_rix

# create a matrix row-wise
matrix(5:10, nrow=2, byrow=TRUE)

# extract third element from second row
mat_rix[2, 3]

# extract second row
mat_rix[2, ]

# now try to extract third column


# extract first and third column
mat_rix[, c(1,3)]

# remove second column
mat_rix[, -2]

# get the number of rows or columns
nrow(mat_rix); ncol(mat_rix)
NROW(mat_rix); NCOL(mat_rix)

# apply nrow() and NROW() on vec_tor

```

*** =solution
```{r}
# create a matrix
mat_rix <- matrix(5:10, nrow=2, ncol=3)

# by default matrices are constructed column-wise
mat_rix

# create a matrix row-wise
matrix(5:10, nrow=2, byrow=TRUE)

# extract third element from second row
mat_rix[2, 3]

# extract second row
mat_rix[2, ]

# now try to extract third column
mat_rix[, 3]

# extract first and third column
mat_rix[, c(1,3)]

# remove second column
mat_rix[, -2]

# get the number of rows or columns
nrow(mat_rix); ncol(mat_rix)
NROW(mat_rix); NCOL(mat_rix)

# apply nrow() and NROW() on vec_tor
nrow(vec_tor)
NROW(vec_tor)
```


*** =sct
```{r}
test_error()
success_msg("So the capitalized functions can be applied to vectors and matrixes")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:9cb7dccabe
## Matrix attributes

Arrays are vectors with a dimension attribute.
Matrices are two-dimensional arrays.
The dimension attribute of a matrix is an integer vector of length 2 (nrow, ncol).
The dimnames attribute is a list, with vector elements containing row and column names.
A named matrix can be subset using row and column names.


*** =instructions
Use attributes(), dim(), rownames(), colnames() and rownames() as instructed.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
mat_rix <- matrix(5:10, nrow=2, ncol=3)
```

*** =sample_code
```{r}
# get matrix attributes
attributes(mat_rix)

# get dimension attribute
dim(mat_rix)

# get class attribute
class(mat_rix)

# rownames and colnames attribute
rownames(mat_rix) <- c("row1", "row2")
colnames(mat_rix) <- c("col1", "col2", "col3")

# use dimnames() to see row and column names


# get the name attributes of mat_rix

```

*** =solution
```{r}
# get matrix attributes
attributes(mat_rix)

# get dimension attribute
dim(mat_rix)

# get class attribute
class(mat_rix)

# rownames and colnames attribute
rownames(mat_rix) <- c("row1", "row2")
colnames(mat_rix) <- c("col1", "col2", "col3")

# use dimnames() to see row and column names
dimnames(mat_rix)

# get the name attributes of mat_rix
names(mat_rix)
```


*** =sct
```{r}
test_error()
success_msg("dimname and name is different!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:cb0ddb5306
## Matrix Subsetting

Matrices can be subset in a similar way as Vectors, either by indices (integers), by characters (names), or boolean vectors.
Subsetting a matrix to a single row or column produces a vector, unless the parameter "drop=FALSE" is used.
Subsetting with the parameter "drop=FALSE" prevents the implicit coercion and preserves the matrix class.
This is an example of implicit coercion in R, which can cause diﬃcult to trace bugs.


*** =instructions
Use "[]" for matrix subsetting, beware of the parameter "drop=FALSE"

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
mat_rix <- matrix(5:10, nrow=2, ncol=3)
rownames(mat_rix) <- c("row1", "row2")
colnames(mat_rix) <- c("col1", "col2", "col3")
```

*** =sample_code
```{r}
# use name of dimension for subsetting
mat_rix["row2", "col3"]

# try to get the subset of only "row1"


# subset columns boolean vector
mat_rix[, c(TRUE, FALSE, TRUE)]

# get subset of row 1 by index
mat_rix[1, ]

# drop=FALSE preserves matrix
mat_rix[1, , drop=FALSE]

# revise the code to create two subsets with and without drop = FALSE argument
mat_1 <- mat_rix[1, ]
mat_2 <- mat_rix[1, ]

# check if the two subsets are matrix using is.matrix()

```

*** =solution
```{r}
# use name of dimension for subsetting
mat_rix["row2", "col3"]

# try to get the subset of only "row1"
mat_rix["row1", ]

# subset columns boolean vector
mat_rix[, c(TRUE, FALSE, TRUE)]

# get subset of row 1 by index
mat_rix[1, ]

# drop=FALSE preserves matrix
mat_rix[1, , drop=FALSE]

# revise the code to create two subsets with and without drop = FALSE argument
mat_1 <- mat_rix[1, ]
mat_2 <- mat_rix[1, , drop=FALSE]

# check if the two subsets are matrix using is.matrix()
is.matrix(mat_1)
is.matrix(mat_2)
```


*** =sct
```{r}
test_error()
success_msg("Add or drop the 'drop = FALSE' argument does make differences!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:5db23f19b0
## Generating Pseudo-Random Numbers

Random number generators produce the same deterministic sequence of numbers after their seed value is reset.
The function set.seed() initializes the random number generator by specifying the seed value.
The function runif() produces random numbers from the uniform distribution.
The function rnorm() produces random numbers from the normal distribution.
The function pnorm() calculates the cumulative normal distribution.
The function qnorm() calculates the inverse cumulative normal distribution.


*** =instructions
Learn to use set.seed() initializers, as well as runif(), rnorm(), pnorm() and qnorm()

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# none
```

*** =sample_code
```{r}
# reset random number generator
set.seed(1121)

# three random numbers from the uniform distribution
runif(3)

# produce another three numbers
runif(3)

# reset random number generator
set.seed(1121)

# produce another three numbers


# produce five random numbers from standard normal distribution using rnorm()


# produce five random numbers from the normal distribution
rnorm(n=5, mean=1, sd=2)

# calculate cumulative standard normal distribution
c(pnorm(-2), pnorm(2))

# calculate inverse cumulative standard normal distribution
c(qnorm(0.75), qnorm(0.25))
```

*** =solution
```{r}
# reset random number generator
set.seed(1121)

# three random numbers from the uniform distribution
runif(3)

# produce another three numbers
runif(3)

# reset random number generator
set.seed(1121)

# produce another three numbers
runif(3)

# produce five random numbers from standard normal distribution using rnorm()
rnorm(5)

# produce five random numbers from the normal distribution
rnorm(n=5, mean=1, sd=2)

# calculate cumulative standard normal distribution
c(pnorm(-2), pnorm(2))

# calculate inverse cumulative standard normal distribution
c(qnorm(0.75), qnorm(0.25))
```


*** =sct
```{r}
test_error()
success_msg("Don't forget to compare output of those runif()")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:7f3b5009e9
## Generating Binomial Random Numbers

A binomial trial is a coin ﬂip, that results in either a success or failure.
The binomial distribution speciﬁes the probability of obtaining a certain number of successes in a sequence of independent binomial trials.
Let p be the probability of obtaining a success in a binomial trial, and let (1−p) be the probability of failure, p = 0.5 corresponds to ﬂipping an unbiased coin.

The function rbinom() produces random numbers from the binomial distribution.

*** =instructions
Learn to use rnorm() and sample() to generate binomial random numbers.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# none
```

*** =sample_code
```{r}
# reset random number generator
set.seed(1121)

# flip unbiased coin once, 20 times
rbinom(n=20, size=1, 0.5)

# number of heads after flipping twice, 20 times


# flip unbiased coin once, 20 times
sample(x=0:1, size=20, replace=TRUE)  # fast
as.numeric(runif(20) < 0.5)  # slower
```

*** =solution
```{r}
# reset random number generator
set.seed(1121)

# flip unbiased coin once, 20 times
rbinom(n=20, size=1, 0.5)

# number of heads after flipping twice, 20 times
rbinom(n=20, size=2, 0.5)

# flip unbiased coin once, 20 times
sample(x=0:1, size=20, replace=TRUE)  # fast
as.numeric(runif(20) < 0.5)  # slower
```


*** =sct
```{r}
test_error()
success_msg("Same results, varied speeds!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:dcee782458
## Generating Random Samples and Permutations

A sample is a subset of elements taken from a set of data elements.
The function sample() produces a random sample form a vector of data elements, By default the size of the sample (the size argument) is equal to the number of elements in the data vector.
So the call sample(data) produces a random permutation of all the elements of data.
If replace=TRUE, then sample() produces samples with replacement.

*** =instructions
Learn to use sample() to generate permutations with or without replacement.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# none
```

*** =sample_code
```{r}
# permutation of five numbers
sample(x=5)

# permutation of four strings
sample(x=c("apple", "grape", "orange", "peach"))

# sample of size three
sample(x=5, size=3)

# try to change the code and sample with replacement with 'replace = TRUE'
sample(x=5)

# sample of strings, change the size to 6
sample(x=c("apple", "grape", "orange", "peach"), size=12, replace=TRUE)
```

*** =solution
```{r}
# permutation of five numbers
sample(x=5)

# permutation of four strings
sample(x=c("apple", "grape", "orange", "peach"))

# sample of size three
sample(x=5, size=3)

# sample with replacement
sample(x=5, replace=TRUE)

# sample of strings
sample(x=c("apple", "grape", "orange", "peach"), size=6, replace=TRUE)
```


*** =sct
```{r}
test_error()
success_msg("Change the 'size' argument to get permutation of differing length")
```



--- type:NormalExercise lang:r xp:100 skills:1 key:d5766325a6
## Statistical Estimators

A data sample is a set of observations of a random variable.
Let {x1,...,xn} be a data sample of a random variable x.
Let x follow a probability distribution with population mean equal to µ and population standard deviation equal to σ.
A statistic is a function of a data sample: f (x1,...,xn).
A statistic is itself a random variable.
A statistical estimator is a statistic that provides an estimate of a distribution parameter.

R has the build-in function to calculate mean `mean()` and standard deviation `sd()`. To calculate the skewness and kurtosis you need to type your own formula.

*** =instructions
Learn to use mean(), sd() to calculate statistical estimators.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# none
```

*** =sample_code
```{r}
# reset random number generator
set.seed(1121)

# sample from Standard Normal Distribution
sam_ple <- rnorm(1000)

# sample mean
mean(sam_ple)

# sample median
median(sam_ple)

# sample standard deviation
sd(sam_ple)

# try to complete the code, save length, mean and sd for higher moments calculation
len <- length()
me_an <- mean()
s_d <- sd()

# calculate the skewness
len / (len - 1) * (len - 2) * sum(((sam_ple - me_an) / s_d) ^ 3)

# calculate the kurtosis
len * (len + 1) / ((len - 1) ^ 3) * sum(((sam_ple - me_an) / s_d) ^ 4)
```

*** =solution
```{r}
# reset random number generator
set.seed(1121)

# sample from Standard Normal Distribution
sam_ple <- rnorm(1000)

# sample mean
mean(sam_ple)

# sample median
median(sam_ple)

# sample standard deviation
sd(sam_ple)

# try to complete the code, save length, mean and sd for higher moments calculation
len <- length(sam_ple)
me_an <- mean(sam_ple)
s_d <- sd(sam_ple)

# calculate the skewness
len / (len - 1) * (len - 2) * sum(((sam_ple - me_an) / s_d) ^ 3)

# calculate the kurtosis
len * (len + 1) / ((len - 1) ^ 3) * sum(((sam_ple - me_an) / s_d) ^ 4)
```


*** =sct
```{r}
test_error()
success_msg("Change the 'size' argument to get permutation of differing length")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:0b1fc2e81d
## Functions in R

R functions have three components: a list of formal arguments, a body containing R code, an environment.
An R function plus its environment is referred to as a function closures, The function body should be enclosed in curly braces {}, unless it contains a single command, then it doesn’t have to enclosed.
The function body doesn’t require a return statement, since by default R functions return the last statement evaluated in the body, args() displays the formal arguments of a function.

*** =instructions
Learn to define your own function in R.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# none
```

*** =sample_code
```{r}
# define a function with two arguments
test_func <- function(first_arg, second_arg) {
  first_arg + second_arg  # returns last evaluated statement
}  # end test_func

# apply the function
test_func(1, 2)

# display argument


# define function that uses variable from enclosure environment
test_func <- function(first_arg, second_arg) {
  first_arg + second_arg + glob_var
}  # end test_func

# create glob_var
glob_var <- 10

# type 3 and 2 as the first, second input of test_func

```

*** =solution
```{r}
# define a function with two arguments
test_func <- function(first_arg, second_arg) {
  first_arg + second_arg  # returns last evaluated statement
}  # end test_func

# apply the function
test_func(1, 2)

# display argument using args() on test_func()
args(test_func)

# define function that uses variable from enclosure environment
test_func <- function(first_arg, second_arg) {
  first_arg + second_arg + glob_var
}  # end test_func

# create glob_var
glob_var <- 10

# type 3 and 2 as the first, second input of test_func
test_func(3, 2)
```


*** =sct
```{r}
test_error()
success_msg("Great! Now what if you redefine test_func() and forget to create glob_var? An error!")
```



--- type:NormalExercise lang:r xp:100 skills:1 key:8d1dbe414e
## Binding Arguments and Default Argument

The formal arguments of a function are deﬁned in its argument list.
When a function is called, it’s passed a list of actual function arguments, Formal arguments can be bound to actual arguments either by name or by position: 
by name: formal arguments are bound to actual arguments with the same name;
by position: the ﬁrst formal argument is bound to the ﬁrst actual argument, etc. 

Binding by name takes precedence over binding by position: ﬁrst all the named arguments are bound, then the remaining arguments are bound by position.
Partial argument names are bound to full names.

Formal arguments may be assigned default values, so that when the actual arguments are missing then their default values are used instead.
Default values are often assigned to function parameters, that determine the function’s behavior.
Default values can be speciﬁed as a vector of strings, representing the possible values of a function’s parameter.
The function match.arg() matches a string to one of the possible values, and returns the matched value, or produces an error if it can’t match it.
The function str() displays the structure of an R object, for example a function name and its formal arguments,


*** =instructions
Learn to bind function arguments, and set default values.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# none
```

*** =sample_code
```{r}
# define a function, with one default value
test_func <- function(first_arg, second_arg, third_arg = 10) {
# last statement of function is return value
  first_arg + 2*second_arg + third_arg
}  # end test_func

# bind by name without defining third argument
test_func(first_arg=3, second_arg=2)

# partial name binding without defining third argument
test_func(first=3, second=2)

# partial name binding, override default value with 12


# bind by position
test_func(3, 2)

# mixed binding
test_func(second_arg=2, third_arg=6, 3)
```

*** =solution
```{r}
# define a function, with one default value
test_func <- function(first_arg, second_arg, third_arg = 10) {
# last statement of function is return value
  first_arg + 2*second_arg + third_arg
}  # end test_func

# bind by name without defining third argument
test_func(first_arg=3, second_arg=2)

# partial name binding without defining third argument
test_func(first=3, second=2)

# partial name binding, override default value with 12
test_func(first=3, second=2, third=12)

# bind by position
test_func(3, 2)

# mixed binding
test_func(second_arg=2, third_arg=6, 3)
```


*** =sct
```{r}
test_error()
success_msg("Enjoy playing the function arguments!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:1c856dfc6f
## Return Values and Invisible Return

The function body doesn’t require a return statement, since by default R functions return the last statement evaluated in the body, return() statements are inserted in logical branches to terminate function execution and return its intended value,

If a return value is wrapped in the function invisible() then the return value isn’t printed.
But if the function is assigned to a variable, then its return value is assigned to that variable, invisible() allows creating functions whose return values can be assigned, but which do not print when they’re not assigned.


*** =instructions
Use return() and invisible() to return value from within a function.

*** =hint
Remember to change the code for the 'return_invisible()' function.

*** =pre_exercise_code
```{r}
# none
```

*** =sample_code
```{r}
# define function that returns NULL for non-numeric argument
test_func <- function(in_put) {
  if (!is.numeric(in_put)) {
    warning(paste("argument", in_put, "isn't numeric"))
    return(NULL)
  }
  2*in_put
}

# test function output with numeric and character input
test_func(2)
test_func("hello")

# change the code to 'invisible()' and define a function that returns invisibly
return_invisible <- function(in_put) {
  return(in_put)
}

# call function with input 2
return_invisible(2)
```

*** =solution
```{r}
# define function that returns NULL for non-numeric argument
test_func <- function(in_put) {
  if (!is.numeric(in_put)) {
    warning(paste("argument", in_put, "isn't numeric"))
    return(NULL)
  }
  2*in_put
}

# test function output with numeric and character input
test_func(2)
test_func("hello")

# change the code to 'invisible()' and define a function that returns invisibly
return_invisible <- function(in_put) {
  invisible(in_put)
}

# call function with input 2
return_invisible(2)
```


*** =sct
```{r}
test_error()
success_msg("Enjoy playing the function arguments!")
```