--- 
title_meta  : Chapter 2
title       : R Objects
description : In this chapter details about R objects will be covered.


--- type:NormalExercise lang:r xp:100 skills:1 key:b1b1305012
## Basic R Objects

The quotation marks "" (or ’’) around a character string tell R that it’s a string, not a variable name.

Vectors are the basic building blocks of R objects, There are no scalars in R, and single values are stored as vectors of length "1".

A character string is also a vector with a single element, with the ﬁrst element of the vector containing the string of text.

The colon binary operator ’:’ produces a vector.
The function c() combines objects into a vector.
The "[1]" symbol means the return value is a vector.
The function is.vector() returns TRUE if its argument is a vector, and returns FALSE otherwise.


*** =instructions
Try to explore different kinds of vector and make use of vector's basic operations.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# create a character vector
vec_char <- "some string"

# check if the single character is a vector


# subset to get the first and second element of the vector


# create a numeric vector
vec_num <- 1:10

# check if the vec_num is a vector


# subset to get the second and third element of vec_num


# now use c() to combine vec_char and vec_num and create vec_c
vec_c <- c(vec_char, vec_num)

# check if vec_c is a vector


# print the second element of vec_c

```

*** =solution
```{r}
# create a character vector
vec_char <- "some string"

# check if the single character is a vector
is.vector(vec_char)

# subset to get the first and second element of the vector
vec_char[1]
vec_char[2]

# create a numeric vector
vec_num <- 1:10

# check if the vec_num is a vector
is.vector(vec_num)

# subset to get the second and third element of vec_num
vec_num[2]
vec_num[3]

# now use c() to combine vec_char and vec_num and create vec_c
vec_c <- c(vec_char, vec_num)

# check if vec_c is a vector
is.vector(vec_c)

# print the second element of vec_c
vec_c[2]
```


*** =sct
```{r}
test_error()
msg <- "Do not remove or change the definition of the variable `vec_num`."
test_object("vec_num", undefined_msg = msg, incorrect_msg = msg)
success_msg("Great!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:05697caf6a
## Character Strings

Character strings are sequences of characters (and vectors of length one), The function nchar() returns the length of a string.

Special characters in strings: 
"\t" for TAB, "\n" for new-line, 
”\\” for a (single) backslash character

The function cat() concatenates strings and echos them to console, without returning any values
The function cat() is useful in user-deﬁned functions.


*** =instructions
Explore characters and strings with functions and basic features. Learn to use `length()` and `nchar()`

*** =hint
Check if object names and function names are correctly typed. Don't miss any variables or functions!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# create a character string
str_var <- "Some string"

# print the entire vector and its second element


# get the length of vector and string


# concatenate and echo to console
cat("Hello", "World!")

# use \n to start a new line when concatenating

```

*** =solution
```{r}
# create a character string
str_var <- "Some string"

# print the entire vector and its second element
str_var
str_var[1]

# get the length of vector and string
length(str_var)
nchar(str_var)

# concatenate and echo to console
cat("Hello", "World!")

# use \n to start a new line when concatenating
cat("Enter\nnewline")
```


*** =sct
```{r}
test_error()
msg <- "Do not remove or change the definition of the variable `str_var`."
test_object("str_var", undefined_msg = msg, incorrect_msg = msg)
success_msg("Excellent!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:b99299073f
## Manipulating Strings

The function paste() concatenates its arguments into a string, coerces them to characters if needed, and returns the string.
If a vector or list is passed to paste(), together with a collapse string, then paste() concatenates the elements into a string, separated by the collapse string.
The function strsplit() splits the elements of a character vector.
Splitting on the "." character requires surrounding it with brackets: "[.]", or using argument fixed=TRUE.
The function substring() extracts or replaces substrings in a character string, The recycling rule extends the length to match the longest object.

*** =instructions
Use functions to manipulate strings. Learn to use `paste()`, `strsplit()`, `substring()`

*** =hint
Check if object names and function names are correctly typed. Don't miss the square bracket!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# create two character string
str_var1 <- "Hello"
str_var2 <- "World!"

# concatenate the two strings and return value


# create a character string and a numeric vector
str_var3 <- "a"
num_var <- 1:4

# convert, recycle and concatenate


# create a new string
hello <- "Hello.World"

# split string on "[.]"


# extract characters from 3 to 6

```

*** =solution
```{r}
# create two character string
str_var1 <- "Hello"
str_var2 <- "World!"

# concatenate the two strings and return value
paste(str_var1, str_var2, sep=' ')

# create a character string and a numeric vector
str_var3 <- "a"
num_var <- 1:4

# convert, recycle and concatenate
paste(str_var3, num_var, sep = ",")

# create a new string
hello <- "Hello.World"

# split string on "[.]"
strsplit(hello, split = "[.]")

# extract characters from 3 to 6
substring(hello, 3, 6)
```


*** =sct
```{r}
test_error()
success_msg("Excellent!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:ab03e239aa
## Regular Expressions in R

R has Regex functions for pattern matching and replacement.

The function gsub() replaces all matches of a pattern in a string.
The function grep() searches for matches of a pattern in a string.
The function glob2rx() converts globbing wildcard patterns into regular expressions.

*** =instructions
Use functions to manipulate strings.

*** =hint
Check if object names and function names are correctly typed. Don't forget to use "i+" instead of "i"

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
my_string <- c("is", "XX", "is this gratis?")
# replace "is" with "XX"


# get indexes of elements in a vector that contains certain string
grep("b+", c("abc", "xyz", "cba d", "bbb"))

# change from the code above to return elements instead of index number


# convert globs into regex (regular expression)
glob2rx("abc.*")

# get the index of elements that contain character "i+" in my_string


# convert my_string to regex

```

*** =solution
```{r}
my_string <- c("is", "XX", "is this gratis?")
# replace "is" with "XX"
gsub("is", "XX", my_string)

# get indexes of elements in a vector that contains certain string
grep("b+", c("abc", "xyz", "cba d", "bbb"))

# change from the code above to return elements instead of index number
grep("b+", c("abc", "xyz", "cba d", "bbb"), value=TRUE)

# convert globs into regex (regular expression)
glob2rx("abc.*")

# get the index of elements that contain character "i+" in my_string
grep("i+", my_string)

# convert my_string to regex
glob2rx(my_string)
```


*** =sct
```{r}
test_error()
success_msg("Remember to look at what's popped up in the console")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:540b8c57a5
## Vectors

Vectors are the basic building blocks of R objects.

For a brief review, we already know that:
There are no scalars in R, and single values are stored as vectors of length "1"
The function c() combines values into a vector
The function is.vector() returns TRUE if its argument is a vector, and returns FALSE otherwise

Beside, we need to knwo that the object letters is a constant and a vector.

To create a vector, R provides several way to do it:
1) The colon operator (":") provides a simple way of creating a numeric vector
2) The function vector() returns a vector of the speciﬁed mode
3) The functions seq(), seq_len(), and seq_along() return a sequence (vector) of numbers
4) The function rep() replicates an object multiple times
5) The functions numeric() and character() return empty (zero length) vectors of the speciﬁed type (not to be confused with a NULL object)
6) Zero length vectors are not NULL objects,


*** =instructions
Use mentioned functions to create vector.

*** =hint
Check if object names and function names are correctly typed. Use "by=" instead of "len=".

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# create vectors from 1 to 10 using ":"
1:10

# create numeric vector of length 10 using `vector()`
vector(mode = "numeric", length = 10)

# create a vector containing five number `10`
rep(10, times = 5)

# create a vector of sequence from 1 to 10
seq(10)

# create a vector of sequence along -5:5
seq(along=(-5:5))

# create a vector of sequence along c("a", "b", "c")
seq_along(c("a", "b", "c"))

# create a vector of decimals from 0 to 1 containing 11 elements
seq(from=0, to=1, len=11)

# create empty character vector of length 5
character(5)

# create empty vector of numerics of length 0
numeric(0)
```

*** =solution
```{r}
# create vectors from 1 to 10 using ":"
1:10

# create numeric vector of length 10 using `vector()`
vector(mode = "numeric", length = 10)

# create a vector containing five number `10`
rep(10, times = 5)

# create a vector of sequence from 1 to 10
seq(10)

# create a vector of sequence along -5:5
seq(along=(-5:5))

# create a vector of sequence along c("a", "b", "c")
seq_along(c("a", "b", "c"))

# create a vector of decimals from 0 to 1 containing 11 elements
seq(from=0, to=1, len=11)

# create empty character vector of length 5
character(5)

# create empty vector of numerics of length 0
numeric(0)
```


*** =sct
```{r}
test_error()
success_msg("See? Empty vector with zero length is not equal to NULL!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:5a2dc65a8a
## Arithmetic and Logical Operations on Vectors

Vectors can be multiplied and squared element by element, as if they were single elements.

When vectors are manipulated as if they were single elements, then R automatically performs a loop over the vector elements, and applies the operation element-wise. This is a very powerful feature of R called vectorized arithmetic. Vectorized arithmetic avoids writing loops and simpliﬁes notation.

Vectors can be combined together and appended.

*** =instructions
Use mentioned functions to conduct vector arithmetic and logical operation.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# multiply the vector 4:8 by 2 without bracket


# multiply the vector 4:8 by 2 with bracket


# divide the vector 4:8 by 2


# create vector
vec_tor <- c(8, 6, 5, 7)

# boolean vector TRUE if element is equal to second one


# boolean vector TRUE for elements greater than six


# append number 2.0 to vector

```

*** =solution
```{r}
# multiply the vector 4:8 by 2 without bracket
2*4:8

# multiply the vector 4:8 by 2 with bracket
2*(4:8)

# divide the vector 4:8 by 2
4:8/2

# create vector
vec_tor <- c(8, 6, 5, 7)

# boolean vector TRUE if element is equal to second one
vec_tor == vec_tor[2]

# boolean vector TRUE for elements greater than six
vec_tor > 6

# append number 2.0 to vector
c(vec_tor, 2.0)
```


*** =sct
```{r}
test_error()
success_msg("Beautiful! Appreciate results in the console!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:2635635720
## Naming and Manipulating Vectors

Vector elements can be assigned names using a list of symbol-value pairs.
The function names() returns the names attribute of an object.
The names attribute of a vector can be modiﬁed by assigning to the names() function (”names() <-” assignment).
The function unname() removes the names attribute.
The function structure() adds attributes (speciﬁed as symbol=value pairs) to an object, and returns it,

*** =instructions
Use names() to get names and assign names. Also learn to use structure() to assign names to a vector.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# create named vector
vec_tor <- c(pi_const=pi, euler=exp(1), gamma=-digamma(1))

# get names of elements


# remove names attribute


# combine two vectors of letters


# create named vector using structure, names are recycled from 'el1' to 'el5'
structure(sample(1:5), names=paste0("el", 1:5))

# now try to assign 1:3 to names of vec_tor


# print out vec_tor

```

*** =solution
```{r}
# create named vector
vec_tor <- c(pi_const=pi, euler=exp(1), gamma=-digamma(1))

# get names of elements
names(vec_tor)

# remove names attribute
unname(vec_tor)

# combine two vectors of letters
c('a', letters[5:10])

# create named vector using structure
structure(sample(1:5), names=paste0("el", 1:5))

# now try to assign 1:3 to names of vec_tor
names(vec_tor) <- 1:3

# print out vec_tor
vec_tor
```


*** =sct
```{r}
test_error()
success_msg("Now you can name a vector with names() and structure() methods!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:e2b1d979f4
## Subsetting Vectors

Vector elements can be subset (indexed, dereferenced) using the "[]" operator.

Vectors can be subset using vectors of:
positive integers, 
negative integers, 
characters (names),
boolean vectors. 

Negative integers remove the vector elements.
Subsetting with zero returns a zero-length vector.
A named vector can be subset using element names.

*** =instructions
Vec_tor has been recreated in this exercise. 
Use "[]" combined with integers, characters and boolean vectors to subset a vector.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
vec_tor <- c(pi_const=pi, euler=exp(1), gamma=-digamma(1))
```

*** =sample_code
```{r}
# extract second element


# extract all elements, except the second element


# extract zero elements - returns zero-length vector


# extract second and third elements by boolean vector


# extract a element using its name "eulery"


# extract multiple elements using a vector of their names


# now create a boolean vector comparing vec_tor to 2, name it as "bool"
bool <- vec_tor > 2

# subset vec_tor with bool to see which element is bigger than 2

```

*** =solution
```{r}
# extract second element
vec_tor[2]

# extract all elements, except the second element
vec_tor[-2]

# extract zero elements - returns zero-length vector
vec_tor[0]

# extract second and third elements by boolean vector
vec_tor[c(FALSE, TRUE, TRUE)]

# extract a element using its name "eulery"
vec_tor["eulery"]

# extract multiple elements using a vector of their names
vec_tor[c("pie", "gammy")]

# now create a boolean vector comparing vec_tor to 2, name it as "bool"
bool <- vec_tor > 2

# subset vec_tor with bool to see which element is bigger than 2
vec_tor[bool]
```


*** =sct
```{r}
test_error()
success_msg("The subsetting provides support for filtering")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:47e9ca9b81
## Filtering Vectors

Filtering means extracting elements from a vector that satisfy a logical condition.
When logical comparison operators are applied to vectors, they produce boolean vectors.
Boolean vectors can then be applied to subset the original vectors, to extract their elements.
The function which() returns the indices of the TRUE elements of a boolean vector or array.


*** =instructions
Vec_tor has been recreated in this exercise. 
Use "[]"  or "which()" to filter vector elements.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
vec_tor <- runif(5)
```

*** =sample_code
```{r}
# first, print the vec_tor


# compare vec_tor to 0.5


# boolean vector of elements equal to the second one


# extract all elements equal to the second one


# extract all elements greater than 0.5


# get index of elements > 0.5


# combine which() and "[]" to extract elements greater than 0.5

```

*** =solution
```{r}
# first, print the vec_tor
vec_tor

# compare vec_tor to 0.5
vec_tor > 0.5

# boolean vector of elements equal to the second one
vec_tor == vec_tor[2]

# extract all elements equal to the second one
vec_tor[vec_tor == vec_tor[2]]

# extract all elements greater than 0.5
vec_tor[vec_tor > 0.5]

# get index of elements > 0.5
which(vec_tor > 0.5)

# combine which() and "[]" to extract elements greater than 0.5
vec_tor[which(vec_tor > 0.5)]
```


*** =sct
```{r}
test_error()
success_msg("Now you can filter a vector with '[]' and which()!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:ab3a8365ff
## Factors

Factors are similar to vectors, but their elements can only take values from a set of levels.
Factors are designed for categorical data which can only take certain values.

The function factor() converts a vector into a factor, Factors have two attributes: class (equal to "factor") and levels (the allowed values).
Although factors aren’t vectors, the data underlying a factor is an integer vector, called an encoding vector.
The function as.numeric() extracts the encoding vector (indices) of a factor.
The function as.vector() coerces a factor to a character vector.

*** =instructions
Create factors with factor(). 
See attributes of factor using attributes() and levels().
Coerce factor using as.numeric() and as.vector().

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
fac_tor <- factor(c('b', 'c', 'd', 'a', 'c', 'b'))
```

*** =sample_code
```{r}
# print the factor vector 'fac_tor'
fac_tor

# subset fac_tor for the third element


# get factor attributes


# get factor levels


# get encoding vector


# coerce factor to character vector

```

*** =solution
```{r}
# print the factor vector 'fac_tor'
fac_tor

# subset fac_tor for the third element
fac_tor[3]

# get factor attributes
attributes(fac_tor)

# get factor levels
levels(fac_tor)

# get encoding vector
as.numeric(fac_tor)

# coerce factor to character vector
as.vector(factor(1:5))
```


*** =sct
```{r}
test_error()
success_msg("See! The numeric vector turned to be a character vector!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:771f2e9f50
## Tables of Categorical Data

The function table() calculates the frequency distribution of categorical data.
A contingency table is a matrix that contains the frequency distribution of variables (factors) contained in a set of data, sapply() applies a function to a vector or a list of objects and returns a vector or a list,


*** =instructions
Use functions and variables as insturcted in the comment.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
fac_tor <- factor(c('b', 'c', 'd', 'a', 'c', 'b'))
```

*** =sample_code
```{r}
# have a look at the factor vector
fac_tor

# get encoding vector


# get unique values


# change the code below to get contingency table
sapply(fac_tor, 
 function(le_vel) {
   sum(fac_tor=le_vel)
 })

# get contingency (frequency) table

```

*** =solution
```{r}
# have a look at the factor vector
fac_tor

# get encoding vector
as.numeric(fac_tor)

# get unique values
unique(fac_tor)

# change the code below to get contingency table using sapply 
sapply(levels(fac_tor), 
 function(le_vel) {
   sum(fac_tor==le_vel)
 })

# get contingency (frequency) table
table(fac_tor)
```


*** =sct
```{r}
test_error()
success_msg("Compared the results you get with sapply() and table()")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:748763cc09
## Classifying Continuous Numeric Data Into Categories

Numeric data that represents a magnitude, intensity, or score can be classiﬁed into categorical data, given a vector of breakpoints.

The breakpoints create intervals that correspond to diﬀerent categories.

The categories combine elements that have a similar numeric magnitude, findInterval() returns the indices of the intervals speciﬁed by "vec" that contain the elements of "x".
If there’s an exact match, then findInterval() returns the same index as function match().
If there’s no exact match, then findInterval() ﬁnds the element of "vec" that is closest to, but not greater than, the element of "x".
If all the elements of "vec" are greater than the element of "x", then findInterval() returns zero,


*** =instructions
Use functions and variables as insturcted in the comment.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# none
```

*** =sample_code
```{r}
# see how we can tweak the outcome findInterval() function with five parameters
str(findInterval)

# get index of the element of "vec" c(3,5,7) that matches 5


# use match()


# no exact match


# use match()


# indices of "vec" that match elements of "x"


# return only indices of inside intervals


# make rightmost interval inclusive

```

*** =solution
```{r}
# see how we can tweak the outcome findInterval() function with five parameters
str(findInterval)

# get index of the element of "vec" c(3,5,7) that matches 5
findInterval(x=5, vec=c(3, 5, 7))

# use match()
match(5, c(3, 5, 7))

# no exact match
findInterval(x=6, vec=c(3, 5, 7))

# use match()
match(6, c(3, 5, 7))

# indices of "vec" that match elements of "x"
findInterval(x=1:8, vec=c(3, 5, 7))

# return only indices of inside intervals
findInterval(x=1:8, vec=c(3, 5, 7), all.inside=TRUE)

# make rightmost interval inclusive
findInterval(x=1:8, vec=c(3, 5, 7), rightmost.closed=TRUE)
```


*** =sct
```{r}
test_error()
success_msg("Congrats! This is the very first step in dividing continuous values")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:91aa35a650
## Classifying Continuous Numeric Data Into Categories, Continued

Temperature can be categorized into ”cold”, ”warm”, ”hot”, etc. 
A named numeric vector of breakpoints can be used to convert a temperature into one of the categories.
Breakpoints correspond to categories of the data.
The ﬁrst breakpoint should correspond to the lowest category, and should have a value less than any of the data.


*** =instructions
Use functions and variables as insturcted in the comment.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# none
```

*** =sample_code
```{r}
# creat named numeric vector of breakpoints
brea_ks <- c("freezing"=0, "very_cold"=30,
       "cold"=50, "pleasant"=60,
       "warm"=80, "hot"=90)

# print out the breakpoints vector
brea_ks

# create a vector of temperatures
tempe_ratures <- runif(10, min=10, max=100)

# change the code to name values according to breakpoints
feels_like <- names(
  brea_ks,findInterval(x=tempe_ratures,
                 vec=brea_ks))

names(tempe_ratures) <- feels_like

# print the categorized vector

```

*** =solution
```{r}
# creat named numeric vector of breakpoints
brea_ks <- c("freezing"=0, "very_cold"=30,
       "cold"=50, "pleasant"=60,
       "warm"=80, "hot"=90)

# print out the breakpoints vector
brea_ks

# create a vector of temperatures
tempe_ratures <- runif(10, min=10, max=100)

# change the code to name values according to breakpoints
feels_like <- names(
  brea_ks[findInterval(x=tempe_ratures,
                 vec=brea_ks)])

names(tempe_ratures) <- feels_like

# print the categorized vector
tempe_ratures
```


*** =sct
```{r}
test_error()
success_msg("Congrats! This is the your first value categorization results!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:ca09ff3670
## Converting Numeric Data Into Factors Using cut()

The function cut() converts a numeric vector into a vector of factors, representing the intervals to which the numeric values belong, cut() divides the range of values into intervals, based on a vector of breaks, cut() then assigns factors to the numeric values, representing the intervals to which the numeric values belong. 
The argument "breaks" is a numeric vector of break points that divide the range of values into intervals.
The argument "labels" is a vector of labels for the intervals.
The argument "right" is a boolean indicating if the intervals should be closed on the right (and open on the left), or vice versa, cut() can produce the same classiﬁcation as findInterval(), but findInterval() is faster than cut(), because it’s a compiled function,

The library microbenchmark is already loaded to the environment

*** =instructions
Use functions and variables as insturcted in the comment.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
library(microbenchmark)
```

*** =sample_code
```{r}
# create a vector of random value
foo <- sample(0:6) + 0.1

# print the vector foo


# use cut to vector
cut(x=foo, breaks=c(2, 4, 6, 8))

# bind the cutted vector with rbind()
rbind(foo, cut(x=foo, breaks=c(2, 4, 6, 8)))

# cut the vector with findInterval()
findInterval(x=1:8, vec=c(2, 4, 6, 8))

# findInterval() is a compiled function, so it is faster than cut()
# change the code to compare their speed
vec_tor <- rnorm(1000)
summary(microbenchmark(
  find_interval=
    findInterval(vec=c(3, 5, 7)),
  cuut=
    cut(breaks=c(3, 5, 7)),
  times=10))[, c(1, 4, 5)]
```

*** =solution
```{r}
# create a vector of random value
foo <- sample(0:6) + 0.1

# print the vector foo
foo

# use cut to vector
cut(x=foo, breaks=c(2, 4, 6, 8))

# bind the cutted vector with rbind()
rbind(foo, cut(x=foo, breaks=c(2, 4, 6, 8)))

# cut the vector with findInterval()
findInterval(x=1:8, vec=c(2, 4, 6, 8))

# findInterval() is a compiled function, so it is faster than cut()
# change the code to compare their speed
vec_tor <- rnorm(1000)
summary(microbenchmark(
  find_interval=
    findInterval(x=vec_tor, vec=c(3, 5, 7)),
  cuut=
    cut(x=vec_tor, breaks=c(3, 5, 7)),
  times=10))[, c(1, 4, 5)]
```


*** =sct
```{r}
test_error()
success_msg("Althogh slower and a little bit different, cut() has similar functionality as findInternval()")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:cd32ce976a
## Matrices

The function matrix() creates a matrix from a vector, and the matrix dimensions.
By default matrix() creates matrices column-wise, unless the argument byrow=TRUE is used.
The elements of matrices can be subset (dereferenced) using the "[]" operator.
The functions nrow() and ncol() return the number of rows and columns of a matrix.
The functions NROW() and NCOL() also return the number of rows or columns of a matrix, but they can also be applied to vectors, and treat vectors as single column matrices.

*** =instructions
Use matrix() to construct matrices, and nrow(), ncol() and NROW(), NCOL().

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
vec_tor <- runif(5)
```

*** =sample_code
```{r}
# create a matrix
mat_rix <- matrix(5:10, nrow=2, ncol=3)

# by default matrices are constructed column-wise
mat_rix

# create a matrix row-wise
matrix(5:10, nrow=2, byrow=TRUE)

# extract third element from second row


# extract second row


# now try to extract third column


# extract first and third column


# remove second column


# get the number of rows or columns
nrow(mat_rix); ncol(mat_rix)
NROW(mat_rix); NCOL(mat_rix)

# apply nrow() and NROW() on vec_tor

```

*** =solution
```{r}
# create a matrix
mat_rix <- matrix(5:10, nrow=2, ncol=3)

# by default matrices are constructed column-wise
mat_rix

# create a matrix row-wise
matrix(5:10, nrow=2, byrow=TRUE)

# extract third element from second row
mat_rix[2, 3]

# extract second row
mat_rix[2, ]

# now try to extract third column
mat_rix[, 3]

# extract first and third column
mat_rix[, c(1,3)]

# remove second column
mat_rix[, -2]

# get the number of rows or columns
nrow(mat_rix); ncol(mat_rix)
NROW(mat_rix); NCOL(mat_rix)

# apply nrow() and NROW() on vec_tor
nrow(vec_tor)
NROW(vec_tor)
```


*** =sct
```{r}
test_error()
success_msg("So the capitalized functions can be applied to vectors and matrixes")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:9cb7dccabe
## Matrix attributes

Arrays are vectors with a dimension attribute.
Matrices are two-dimensional arrays.
The dimension attribute of a matrix is an integer vector of length 2 (nrow, ncol).
The dimnames attribute is a list, with vector elements containing row and column names.
A named matrix can be subset using row and column names.


*** =instructions
Use attributes(), dim(), rownames(), colnames() and rownames() as instructed.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
mat_rix <- matrix(5:10, nrow=2, ncol=3)
```

*** =sample_code
```{r}
# get matrix attributes


# get dimension attribute


# get class attribute


# change the code to assign rownames and colnames attribute
c("row1", "row2")
c("col1", "col2", "col3")

# use dimnames() to see row and column names


# get the name attributes of mat_rix

```

*** =solution
```{r}
# get matrix attributes
attributes(mat_rix)

# get dimension attribute
dim(mat_rix)

# get class attribute
class(mat_rix)

# rownames and colnames attribute
rownames(mat_rix) <- c("row1", "row2")
colnames(mat_rix) <- c("col1", "col2", "col3")

# use dimnames() to see row and column names
dimnames(mat_rix)

# get the name attributes of mat_rix
names(mat_rix)
```


*** =sct
```{r}
test_error()
success_msg("dimname and name is different!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:cb0ddb5306
## Matrix Subsetting

Matrices can be subset in a similar way as Vectors, either by indices (integers), by characters (names), or boolean vectors.
Subsetting a matrix to a single row or column produces a vector, unless the parameter "drop=FALSE" is used.
Subsetting with the parameter "drop=FALSE" prevents the implicit coercion and preserves the matrix class.
This is an example of implicit coercion in R, which can cause diﬃcult to trace bugs.


*** =instructions
Use "[]" for matrix subsetting, beware of the parameter "drop=FALSE"

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
mat_rix <- matrix(5:10, nrow=2, ncol=3)
rownames(mat_rix) <- c("row1", "row2")
colnames(mat_rix) <- c("col1", "col2", "col3")
```

*** =sample_code
```{r}
# subset column1 by name 
mat_rix[, "col1"]

# subset columns 1,3 by boolean vector
mat_rix[, c(TRUE, FALSE, TRUE)]

# get subset of row 1 by index
mat_rix[1, ]

# drop=FALSE preserves matrix
mat_rix[1, , drop=FALSE]

# revise the code to create two subsets with and without drop = FALSE argument
mat_1 <- mat_rix[1, ]
mat_2 <- mat_rix[1, ]

# check if the two subsets are matrix using is.matrix()

```

*** =solution
```{r}
# subset column1 by name 
mat_rix[, "col1"]

# subset columns 1,3 by boolean vector
mat_rix[, c(TRUE, FALSE, TRUE)]

# get subset of row 1 by index
mat_rix[1, ]

# drop=FALSE preserves matrix
mat_rix[1, , drop=FALSE]

# revise the code to create two subsets with and without drop = FALSE argument
mat_1 <- mat_rix[1, ]
mat_2 <- mat_rix[1, , drop=FALSE]

# check if the two subsets are matrix using is.matrix()
is.matrix(mat_1)
is.matrix(mat_2)
```


*** =sct
```{r}
test_error()
success_msg("Add or drop the 'drop = FALSE' argument does make differences!")
```

