--- 
title_meta  : Chapter 1
title       : R Environment
description : "This is the first chapter in the Introduction to R course. You will get to know the R environment. You will also learn basic manipulations on R objects, which is essential for R. Let's get started!"
attachments : 





--- type:NormalExercise lang:r xp:100 skills:1 key:5714863ba8
## Variables assignment

In the script window on the upper right you should type the *R* code required for the exercises. When you hit the *Submit Answer* button, your *R* code is executed, and you get a message indicating whether your *R* code is correct. The output of your submission is shown in the *R* console on the lower right.

You can also execute *R* commands in the console on the lower right. When you type in the console, your submission will not be checked for correctness! Try, for example, to type in `2 + 1` and hit Enter. *R* should return `[1] 3`.

Variables are created by an assignment operation, and they don’t have to be declared, The standard assignment operator in R is the arrow symbol `<-`, R interprets text in quotes (`””`) as character strings, Text that is not in quotes (`””`) is interpreted as a symbol or expression, Typing a symbol or expression evaluates it.

*** =instructions
In the script window on the upper right, assign number 2 and a string "Hello World!" to variables, and hit the *Submit Answer* button.

*** =hint
The variable assignments are already presented, just type the variable name and check results. Remember not to add quotes to them!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# assign number to a variable
myvariable1 <- 2

# assign a string to a variable
myvariable2 <- "Hello World!"

# now write R code of variable names to check if the assignments are successfull
```

*** =solution
```{r}
# assign number to a variable
myvariable1 <- 2

# assign a string to a variable
myvariable2 <- "Hello World!"

# now write R code to check if the assignments are successfull
myvariable1
myvariable2

```

*** =sct
```{r}
success_msg("Awesome! See how the console shows the result of the R code you submitted?")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:ab37530088
## Inserting comments into your code

*R* uses the hash `#` sign to mark text as comments.  
All text after the hash `#` sign is treated as a comment, and is not executed as code.  
For example, the text *Calculate 3 + 4* is a comment and is ignored during execution.

*** =instructions
Add a line with the comment mark *#* and  text *myvariable <- 1*, in the script window.

*** =hint
Simply add the line `# add comments to myvariable <- 1` above the *R* code that assigns values *myvariable1 <- 2*.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
myvariable <- 10

# assign number to a variable
myvariable1 <- 2

# assign a string to a variable
myvariable2 <- "Hello World!"

# add comments mark before assign 1 to 'myvariable1'


# now write R code to check if the assignments are successfull
```

*** =solution
```{r}
myvariable <- 10

# assign number to a variable
myvariable1 <- 2

# assign a string to a variable
myvariable2 <- "Hello World!"

# add comments
# myvariable <- 1

# now write R code to check if the assignments are successfull
myvariable
myvariable1
myvariable2
```

*** =sct
```{r}
success_msg("Awesome! See? That assignment in the comment does not take effect.")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:93f879e02e
## Explore sessions in R

R session has many default settings, which you can learn to configure or make use of.

The function getwd() returns a vector of length 1, with the ﬁrst element containing a string with the name of the current working directory (cwd).
The function setwd() accepts a character string as input (the name of the directory), and sets the working directory to that string. 
With Sys.time() you can get system time with time zone of the R session; or you can just get system date with Sys.Date().

*** =instructions
Use the function getwd(), Sys.time() and Sys.Date() to get strings of working directory, system time and system date.

*** =hint
Simply the function name, copy & paste.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# get string with the name of the current working directory


# get current system time with time zone


# get current date

```

*** =solution
```{r}
# get string with the name of the current working directory
getwd()

# get current system time with time zone
Sys.time()

# get current date
Sys.Date()
```


*** =sct
```{r}
success_msg("Great! Look at the results, play with them if you would like to")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:b1bb4402e2
## More about R sessions

The function sessionInfo() returns information about the current R session:
R version;
OS platform;
locale settings;
list of packages that are loaded and attached to the search path;
list of packages that are loaded, but not attached to the search path


*** =instructions
Use the function sessionInfo() to get detailed information about current R session.

*** =hint
Simply the function name, copy & paste.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# get information about the current R session

```

*** =solution
```{r}
# get information about the current R session
sessionInfo()
```


*** =sct
```{r}
success_msg("Correct! Remeber to see what's pop up in the console")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:2f547ffda2
## Environment Variables

R uses environment variables to store information about its environment, such as paths to directories containing ﬁles used by R (startup, history, OS).
For example the environment variables: 
R USER and HOME store the R user Home directory, 
R HOME stores the root directory of the R installation, 
The functions Sys.getenv() and Sys.setenv() display and set the values environment variables, 
Sys.getenv("env var") displays the environment variable "env var", 
Sys.setenv("env var=value") sets the environment variable "env var" equal to "value",


*** =instructions
Use `Sys.getenv()` to display values of environment variables.
/nChange parameters as instructed and see the results.

*** =hint
Simply copy the function name and change parameters inside.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# get the subset from 5th to 7th element for list of environment variable by subsetting
Sys.getenv()[5:7]

# get user HOME directory 


# get R_HOME directory 


# get "etc" sub-directory of R_HOME

```

*** =solution
```{r}
# get the subset from 5th to 7th element for list of environment variable by subsetting
Sys.getenv()[5:7]

# get user HOME directory 
Sys.getenv("Home")

# get R_HOME directory 
Sys.getenv("R_home")

# get "etc" sub-directory of R_HOME
Sys.getenv("env_var")
```


*** =sct
```{r}
success_msg("Correct! Remeber to see what's pop up in the console")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:4d04bf29c2
## Global Options Settings

R uses a list of global options which aﬀect how R computes and displays results./n
The function options() either sets or displays the values of global options:/n 
  options("globop") displays the current value of option "globop";/n
  getOption("globop") displays the current value of option "globop";/n
  options(globop=value) sets the option "globop" equal to "value";


*** =instructions
Change parameters as instructed and see the results.

*** =hint
Remeber to capitalize TRUE and don't use commas for the last code!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# get number of digits printed for numeric values
options("digits")

# get the current value of "digits"


# set the digits option to 3, type 0.12345 and see results

```

*** =solution
```{r}
# get number of digits printed for numeric values
getOption("digits")

# get the current value of "digits"
options("digits")

# set the digits option to 3, type 0.12345 and see results
options(digits = 3)
0.12345
```


*** =sct
```{r}
success_msg("OK, now see what's pop up in the console")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:58ca8cb732
## The R Workspace

The workspace is the current R working environment, which includes all user-deﬁned objects and the command history./n
The function ls() returns names of objects in the R workspace./n
The function rm() removes objects from the R workspace./n
The workspace can be saved into and loaded back from an *.RData ﬁle (binary ﬁle format)./n
The function save.image() saves the whole workspace. /n
The function save() saves just the selected objects./n
The function load() reads data from *.RData ﬁles, and invisibly returns a vector of names of objects created in the workspace./n

When you quit R you’ll be prompted ”Save workspace image?” If you answer YES then the workspace will be saved into the .RData ﬁle in the cwd./n
When you start R again, the workspace will be automatically loaded from the existing .RData ﬁle./n
/n
The function history() displays recent commands, You can also save and load the command history from a ﬁle./n

*** =instructions
Use the function ls() and rm() to see all objects in the environment and remove some of them.

*** =hint
Simply the function name, copy & paste as instructed by comments.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
variable_1 <- 5
variable_2 <- "sand"

# see what variable is currently in the workspace


# remove all elements in the workspace
rm(list = ls())

# check what is left in the workspace

```

*** =solution
```{r}
variable_1 <- 5
variable_2 <- "sand"
# see what variable is currently in the workspace
ls()

# remove all elements in the workspace
rm(list = ls())

# check what is left in the workspace
ls()
```


*** =sct
```{r}
success_msg("Great! Look at the results, play with them if you would like to")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:39d8f62180
## Data Objects in R

All data objects in R are vectors, or consist of vectors.
Single numbers and character strings are vectors of length "1".
Atomic vectors are homogeneous objects whose elements are all of the same mode (type).
Lists and data frames are recursive (heterogeneous) objects, whose elements can be vectors of diﬀerent mode.
The functions is.atomic() and is.recursive() return logical values depending on whether their arguments are atomic or recursive.

*** =instructions
Use function is.atomic() and is.recursive() to check whether the object is atomic or recursive.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# define objects of single number, vector, list and dataframe
v1 <- 1
v2 <- 1:10
v3 <- list(aa=c('a', 'b'), bb=1:5)
v4 <- data.frame(aa=c('a', 'b'), bb=1:2)

# find out if v1, v2 is atomic or recursive
is.atomic(v1)
is.recursive(v1)

# find out if v2 (vector) is atomic or recursive


# use similar process to find out if v3 (list) is atomic or recursive


# do the same to v4 (dataframe)

```

*** =solution
```{r}
# define objects of single number, vector, list and dataframe
v1 <- 1
v2 <- 1:10
v3 <- list(aa=c('a', 'b'), bb=1:5)
v4 <- data.frame(aa=c('a', 'b'), bb=1:2)

# find out if v1, v2 is atomic or recursive
is.atomic(v1)
is.recursive(v1)

# find out if v2 (vector) is atomic or recursive
is.atomic(v2)
is.recursive(v2)

# use similar process to find out if v3 (list) is atomic or recursive
is.atomic(v3)
is.recursive(v3)

# do the same to v4 (dataframe)
is.atomic(v4)
is.recursive(v4)
```


*** =sct
```{r}
success_msg("You can scroll down and console and see the output")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:a5293e3bde
## Type, Mode, and Class of Objects

The type, mode, and class are character strings representing various object properties.
The type of an atomic object represents how it’s stored in memory (logical, character, integer, double, etc.)
The mode of an atomic object is the kind of data it represents (logical, character, numeric, etc.)
The mode of an object often coincides with its type (except for integer and double types)
Recursive objects (lists, data frames) have both type and mode equal to the recursive type (list)
The class of an object is given by either an explicit class attribute, or is derived from the object mode and its dim attribute (implicit class)
The function class() returns the explicit or implicit class of an object.
The object class is used for method dispatching in the S3 object-oriented programming system in R

*** =instructions
Use function typeof(), mode() and class() to explore object type, mode and class.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# define objects of vector, matrix, list and dataframe
v1 <- 1:10
v2 <- matrix(1:10, 2, 5)
v3 <- list(aa=c('a', 'b'), bb=1:5)
v4 <- data.frame(aa=c('a', 'b'), bb=1:2)

# find out the type, mode and class of vector and matrix
c(typeof(v1), mode(v1), class(v1))
c(typeof(v2), mode(v2), class(v2))

# use the same code to check list and dataframe

```

*** =solution
```{r}
# define objects of vector, matrix, list and dataframe
v1 <- 1:10
v2 <- matrix(1:10, 2, 5)
v3 <- list(aa=c('a', 'b'), bb=1:5)
v4 <- data.frame(aa=c('a', 'b'), bb=1:2)

# find out the type, mode and class of vector and matrix
c(typeof(v1), mode(v1), class(v1))
c(typeof(v2), mode(v2), class(v2))

# now use the same code to check list and dataframe
c(typeof(v3), mode(v3), class(v3))
c(typeof(v4), mode(v4), class(v4))
```


*** =sct
```{r}
success_msg("Good Job!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:85b773db52
## R Object Attributes

R objects can have diﬀerent attributes, such as: names, dimnames, dimensions, class, etc.
The attributes of an object is a named list of symbol=value pairs.
The function attributes() returns the attributes of an object.
The attributes of an R object can be modiﬁed using the "attributes()" assignment.

The function structure() adds attributes (speciﬁed as symbol=value pairs) to an object, and returns it.
A vector that is assigned an attribute other than names is not treated as a vector.
The function is.vector() returns TRUE if its argument is a vector, and returns FALSE otherwise.


*** =instructions
Use function attributes(), structure() and is.vector() to explore object attributes.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# check the attributes of a simple vector 1:10


my_var <- c(pi=pi, euler=exp(1), gamma=-digamma(1))
# check the attributes of names vector `my_var`


# modify the attributes of an object, convert the class to "Date"


# print the object and see what's changed


# use structure to create a vector of Date from 1:10


my_list <- list(1, 2)
# now use attributes assignment method to change the attributes of my_list to a vector of Dates


# print out your result


# check if it is a vector

```

*** =solution
```{r}
# check the attributes of a simple vector 1:10
attributes(1:10)

my_var <- c(pi=pi, euler=exp(1), gamma=-digamma(1))
# check the attributes of names vector `my_var`
attributes(my_var)

# modify the attributes of an object, convert the class to "Date"
attributes(my_var) <- list(class = "Date")

# print the object and see what's changed
my_var

# use structure to create a vector of Date from 1:10
structure(1:10, class = "Date")

my_list <- list(1, 2)
# now use attributes assignment method to change the attributes of my_list to a vector of Dates
attributes(my_list) <- list(class = "Date")

# print out your result
my_list

# check if it is a vector
is.vector(my_list)
```


*** =sct
```{r}
test_error()
msg <- "Do not remove or change the definition of the variable `my_list`."
test_object("my_list", undefined_msg = msg, incorrect_msg = msg)
test_function("is.vector", "x",
              not_called_msg = "Make sure to call the function [`is.vector()`] to see if the object is a vector.",
              incorrect_msg = "Have you passed the correct variable to the function [`is.vector()`]?")
success_msg("Good Job!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:792449dde3
## Implicit Class of Objects

We will revisit the class of objects in this exercise. This time there are more details:

If an object has no explicit class, then its implicit class is derived from its mode and dim attribute (except for integer vectors which have the implicit class "integer" derived from their type).

If an atomic object has a dim attribute, then its implicit class is matrix or array.
Data frames have no explicit dim attribute, but dim() returns a value, so they have an implicit dim attribute.


*** =instructions
Now that a dataframe is created. Use function typeof(), mode(), class() and attributes() to explore results.

*** =hint
Check if object names and function names are correctly typed. Remember to use all of the functions in the instructions.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# integer implicit class derived from type
my_var_1 <- vector(mode="integer", length=10)

# check the type, mode and class of my_var1


# numeric implicit class derived from mode
my_var_2 <- vector(mode="numeric", length=10)

# check the type, mode and class of my_var2


# add c(5,2) to dim attribute of `my_var_1` changes implicit class to matrix


# check the type, mode and class of my_var1


# define a new dataframe
my_var <- data.frame(aa=c('a', 'b'), bb=1:2)

# now explore the type, class, mode of a data frame


# check the attributes of my_var

```

*** =solution
```{r}
# integer implicit class derived from type
my_var_1 <- vector(mode="integer", length=10)

# check the type, mode and class of my_var1
c(typeof(my_var_1), mode(my_var_1), class(my_var_1))

# numeric implicit class derived from mode
my_var_2 <- vector(mode="numeric", length=10)

# check the type, mode and class of my_var2
c(typeof(my_var_2), mode(my_var_2), class(my_var_2))

# add c(5,2) to dim attribute of `my_var_1` changes implicit class to matrix
dim(my_var_1) <- c(5, 2)

# check the type, mode and class of my_var1
c(typeof(my_var_1), mode(my_var_1), class(my_var_1))

# define a new dataframe
my_var <- data.frame(aa=c('a', 'b'), bb=1:2)

# now explore the type, class, mode of a data frame
c(typeof(my_var), mode(my_var), class(my_var))

# check the attributes of my_var
attributes(my_var)
```


*** =sct
```{r}
test_error()
msg <- "Do not remove or change the definition of the variable `my_var`."
test_object("my_var", undefined_msg = msg, incorrect_msg = msg)
#test_function("typeof", "x",
#              not_called_msg = "Make sure to call the function [`typeof()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`typeof()`]?")
#test_function("mode", "x",
#              not_called_msg = "Make sure to call the function [`mode()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`mode()`]?")
#test_function("class", "x",
#              not_called_msg = "Make sure to call the function [`class()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`class()`]?")
#test_function("attributes", "x",
#              not_called_msg = "Make sure to call the function [`attributes()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`attributes()`]?")
success_msg("Good Job!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:914dcdc4c2
## Object Coercion

Coercion means changing the type, mode, or class of an object, often without changing the underlying data. 
Changing the mode of an object can change its class as well, but not always.
Objects can be explicitly coerced using the "as.*" coercion functions.
Most coercion functions strip the attributes from the object.
Implicit coercion occurs when objects with diﬀerent modes are combined into a vector, forcing the elements to have the same mode.
Implicit coercion can cause bugs that are diﬃcult to trace.
The rule is that coercion is into larger types (numeric objects are coerced into character strings).

Notice: Coercion can introduce bad data, such as NA values.


*** =instructions
Use coercison to convert object types with functions of `as.*`.

*** =hint
Check if object names and function names are correctly typed.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# create a numeric vector
my_var_1 <- vector(mode="integer", length=10)

# check the type, mode and class of my_var_1


# coerce the vector into character


# check the type, mode and class of my_var_1 after the change


# create a mixed vector
my_var <- c(2, "a", 3.2)

# check the type, mode and class of my_var


# now coerce the vector into integer


# check the type, mode and class of my_var

```

*** =solution
```{r}
# create a numeric vector
my_var_1 <- vector(mode="integer", length=10)

# check the type, mode and class of my_var_1
c(typeof(my_var_1), mode(my_var_1), class(my_var_1))

# coerce the vector into character
my_var_1 <- as.character(my_var_1)

# check the type, mode and class of my_var_1 after the change
c(typeof(my_var_1), mode(my_var_1), class(my_var_1))

# create a mixed vector
my_var <- c(2, "a", 3.2)

# check the type, mode and class of my_var
c(typeof(my_var), mode(my_var), class(my_var))

# now coerce the vector into integer
my_var <- as.integer(my_var)

# check the type, mode and class of my_var
c(typeof(my_var), mode(my_var), class(my_var))
```


*** =sct
```{r}
test_error()
msg <- "Do not remove or change the definition of the variable `my_var`."
test_object("my_var", undefined_msg = msg, incorrect_msg = msg)
#test_function("typeof", "x",
#              not_called_msg = "Make sure to call the function [`typeof()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`typeof()`]?")
#test_function("mode", "x",
#              not_called_msg = "Make sure to call the function [`mode()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`mode()`]?")
#test_function("class", "x",
#              not_called_msg = "Make sure to call the function [`class()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`class()`]?")
#test_function("attributes", "x",
#              not_called_msg = "Make sure to call the function [`attributes()`].",
#              incorrect_msg = "Have you passed the correct variable to the function [`attributes()`]?")
success_msg("Good Job!")
```

